---
title: "Databases and Datasets"
author: "K Arnold"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: FALSE
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
source("../slides-common.R")
slideSetup()
knitr::opts_chunk$set(echo = TRUE)
```

## Data Formats

* Tabular, delimited (CSV, TSV) or fixed-width
* Tabular, structured: Excel, SPSS/Stata/SAS, etc.
* Hierarchical
  * JSON
  * XML
  * HTML
* Database
  * SQLite (most apps)
  * PostgreSQL / MySQL / Oracle / Microsoft SQL Server
  * Google BigQuery

(For a big table, see the [`rio` package vignette](https://cloud.r-project.org/web/packages/rio/vignettes/rio.html))

---

## Spreadsheets vs Databases

.pull-left[
* Often exchanged by email
* Hope the format doesn't change
* [Capacity limited](https://www.bbc.com/news/technology-54423988) ([other article](https://theconversation.com/why-you-should-never-use-microsoft-excel-to-count-coronavirus-cases-147681))
* Slow to query
* Decentralized (resilient?)
]

.pull-right[
* Centralized, highly available servers
* Documented schema
* Large capacity
* Fast queries
]


---

## Grammar of Data

| dplyr (`data %>% `) | Pandas | SQL |
|-------|--------|-----|
| `select(col1, col2)` | `data[['col1', 'col2']]` | `SELECT col1, col2 FROM data` |
| `filter(col1 > 5)` | `data.query('col1 > 5')` | `SELECT * FROM data WHERE col1 > 5` |
| `left_join(data2, by = "col2")` | `pd.merge(data, data2, by="col2", type="left")` | `SELECT * FROM data LEFT JOIN data2 ON data.col2 == data2.col2` |
| `group_by(col2) %>% summarize(m = max(col1))` | `data.group_by('col2')['col1'].max()` | `SELECT max(col1) AS m FROM data GROUP BY col2` |
| `pivot_longer()` | `data.melt()` or `data.pivot()` | **No standard approach** |

* [SQLite syntax](https://sqlite.org/lang_select.html)
* [More examples](https://pandas.pydata.org/pandas-docs/stable/getting_started/comparison/comparison_with_r.html)

---

## BigQuery

```{r}
billing_project <- "calvindsdev"
```

Several public datasets are available:

```{r}
bq_connection <- DBI::dbConnect(
  bigrquery::bigquery(),
  project = "bigquery-public-data",
  billing = billing_project
)

covid_open_data <- 
  bq_connection %>% tbl("bigquery-public-data.covid19_open_data.covid19_open_data")
covid_open_data %>% glimpse()
```

```{r}
world_bank_global_pop <- 
  bq_connection %>% 
  tbl("bigquery-public-data.world_bank_global_population.population_by_country")
world_bank_global_pop %>% glimpse()
```


```{r}
covid_open_data %>%
  filter(date == '2020-11-13', aggregation_level == 0) %>% 
  left_join(
  world_bank_global_pop %>% select(iso_3166_1_alpha_3 = country_code, population_2018 = year_2018),
  by = c("iso_3166_1_alpha_3")
) %>%
  filter(population_2018 > 100000000) %>% 
  arrange(desc(cumulative_confirmed)) %>% 
  head(10)
```


<https://console.cloud.google.com/marketplace/product/bigquery-public-datasets/covid19-open-data?filter=solution-type:dataset&q=covid19&project=calvindsdev>

```
WITH
  country_pop AS (
  SELECT
    country_code AS iso_3166_1_alpha_3,
    year_2018 AS population_2018
  FROM
    `bigquery-public-data.world_bank_global_population.population_by_country`)
SELECT
  country_code,
  country_name,
  cumulative_confirmed AS june_confirmed_cases,
  population_2018,
  ROUND(cumulative_confirmed/population_2018 * 100,2) AS case_percent
FROM
  `bigquery-public-data.covid19_open_data.covid19_open_data`
JOIN
  country_pop
USING
  (iso_3166_1_alpha_3)
WHERE
  date = '2020-11-13'
  AND aggregation_level = 0
  AND population_2018 > 100000000
ORDER BY
  case_percent DESC
```

See also: <https://www.reddit.com/r/bigquery/comments/3cej2b/17_billion_reddit_comments_loaded_on_bigquery/>
from <https://fivethirtyeight.com/features/dissecting-trumps-most-rabid-online-following/>
