---
title: "Midterm grading notes"
author: "K Arnold"
date: "11/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Overall

* Many people struggled with counting the number of failed joins.
  * A failed join is indicated by missing data (`is.na`).
  * You can focus only on parts that have missing data by using `filter` (remember the Nobel lab?)
  * `distinct()` and `nrow`, don't try to count by hand.
  

### To me

* I should not have taught `count`.


## Car prices and eco

### Question 11

* Used `count` instead of `filter` (-2)



## Loading responses

```{r}
response_dir <- "../fa20/midterm/responses"
response_files <- withr::with_dir(response_dir, {
  Sys.glob("*/*/*")})
```

```{r}
read_file(file.path(response_dir, "Q10 - doors-engine-size-repro/1355970 - Adam DenHaan - Attempt1 - 2020-11-02 6_28 pm/Attempt1_textresponse"))
```


```{r}
responses <- tibble(filename = response_files) %>% 
  separate(filename, into = c("question", "attempt", "text_filename"), sep = "/", remove = FALSE) %>% 
  separate(attempt, into = c("id", "student", "attempt_num", "time"), sep = "-", extra = "merge") %>% 
  mutate(attempt_num = as.numeric(str_replace(attempt_num, "Attempt", ""))) %>% 
  group_by(question, student) %>% 
  arrange(attempt_num) %>% 
  slice_tail(n = 1) %>% 
  mutate(contents = read_file(file.path(response_dir, filename))) %>% 
  relocate(filename, .after = everything())
```

```{r}
# https://stackoverflow.com/a/61647053/69707
mixedrank = function(x) order(gtools::mixedorder(x))
responses <- responses %>% 
  arrange(mixedrank(question))
```

## Show all results

```{r results='asis'}
responses %>% 
  select(student, question, contents) %>% 
  mutate(md = glue::glue("#### {student}\n\n```r\n{contents}\n```\n")) %>% 
  group_by(question) %>% 
  arrange(contents) %>% 
  summarize(md = paste0("\n### ", question[[1]], "\n\n", paste0(md, collapse = '\n\n'))) %>% 
  ungroup() %>% 
  arrange(mixedrank(question)) %>% 
  summarize(md = paste0(md, collapse='\n')) %>% 
  pull(md) %>% cat()
```


## Unfortunately I caught a cheat...

```{r results='asis'}
responses %>% 
  filter(str_detect(student, "Hyein") | str_detect(student, "Krista")) %>% 
  select(student, question, contents) %>% 
  mutate(md = glue::glue("#### {student}\n\n```r\n{contents}\n```\n")) %>% 
  group_by(question) %>% 
  summarize(md = paste0("\n### ", question[[1]], "\n\n", paste0(md, collapse = '\n\n'))) %>% 
  ungroup() %>% 
  arrange(mixedrank(question)) %>% 
  summarize(md = paste0(md, collapse='\n')) %>% 
  pull(md) %>% cat()
```


