---
title: "Midterm grading notes"
author: "K Arnold"
date: "11/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Overall

* Many people struggled with counting the number of failed joins.
  * A failed join is indicated by missing data (`is.na`).
  * You can focus only on parts that have missing data by using `filter` (remember the Nobel lab?)
  * `distinct()` and `nrow`, don't try to count by hand.
  

### To me

* I should not have taught `count`.


## Car prices and eco

### Question 11

* Used `count` instead of `filter` (-2)



## Loading responses

```{r}
response_dir <- "../fa20/midterm/responses"
response_files <- withr::with_dir(response_dir, {
  Sys.glob("*/*/*")})
```

```{r}
read_file(file.path(response_dir, "Q10 - doors-engine-size-repro/1355970 - Adam DenHaan - Attempt1 - 2020-11-02 6_28 pm/Attempt1_textresponse"))
```


```{r}
# https://stackoverflow.com/a/61647053/69707
mixedrank = function(x) order(gtools::mixedorder(x))
responses <- responses %>% 
  arrange(mixedrank(question))
```


```{r}
responses <- tibble(filename = response_files) %>% 
  separate(filename, into = c("question", "attempt", "text_filename"), sep = "/", remove = FALSE) %>% 
  separate(attempt, into = c("id", "student", "attempt_num", "time"), sep = "-", extra = "merge") %>% 
  mutate(attempt_num = as.numeric(str_replace(attempt_num, "Attempt", ""))) %>% 
  mutate(student = str_trim(student)) %>% 
  group_by(question, student) %>% 
  arrange(attempt_num) %>% 
  slice_tail(n = 1) %>% 
  mutate(contents = read_file(file.path(response_dir, filename))) %>% 
  relocate(filename, .after = everything())
```

## Show all results

```{r}
summarizeFuncCalls <- function(text) {
  parsed <- parse(text = text, keep.source = TRUE)
  parseData <- utils::getParseData(parsed)
  if (!("token" %in% colnames(parseData))) {
    cat(as.character(parseData), file=stderr())
    stop()
  }
  texts <- filter(parseData, token == "SYMBOL_FUNCTION_CALL")$text
  paste0(texts, collapse = ",")
}
```

```{r}
#responses %>% arrange(student) %>% ungroup() %>%  distinct(student) %>% clipr::write_clip()
```

```{r}
responses %>% 
  ungroup() %>% 
  complete(student, question, fill = list(contents="MISSING")) %>% 
  filter(contents == "MISSING")
```

```{r results='asis'}
dir.create("out")
responses %>% 
  ungroup() %>% 
  bind_rows(tibble(student = "template", question = unique(responses$question))) %>% 
  complete(student, question, fill = list(contents="MISSING")) %>% 
  group_by(student) %>% 
  arrange(mixedrank(question)) %>% 
  summarise(md = str_c(
    paste0("# ", student[[1]], "\n\\newpage\n"),
    glue::glue("## {question}\n\n```r\n{contents}\n```\n\\newpage") %>% paste0(collapse = "\n")
  )) %>% 
  rowwise() %>% 
  mutate(
    filename = glue::glue("out/{student}.md"),
    pdf_filename = glue::glue("out/{student}.pdf"),
    res = write_lines(md, filename),
    res2 = sys::exec_wait("pandoc", c("-V", "geometry:landscape", "-f", "markdown", "-t", "pdf", filename, "-o", pdf_filename)))

```


```{r results='asis'}
responses %>% 
  ungroup() %>% 
  mutate(func_calls = map_chr(contents, possibly(summarizeFuncCalls, "PARSE_ERROR", quiet = TRUE))) %>% 
  select(student, question, func_calls, contents) %>% 
  mutate(md = glue::glue("#### {student}\n\n{func_calls}\n```r\n{contents}\n```\n")) %>% 
  select(student, question, md, func_calls) %>% 
  group_by(question) %>% 
  arrange(func_calls) %>% 
  summarize(md = paste0("\n### ", question[[1]], "\n\n", paste0(md, collapse = '\n\n'))) %>% 
  ungroup() %>% 
  arrange(mixedrank(question)) %>% 
  summarize(md = paste0(md, collapse='\n')) %>% 
  pull(md) %>% cat()
```




## Unfortunately I caught a cheat...

```{r results='asis', eval=FALSE}
responses %>% 
  filter(str_detect(student, "Hyein") | str_detect(student, "Krista")) %>% 
  select(student, question, contents) %>% 
  mutate(md = glue::glue("#### {student}\n\n```r\n{contents}\n```\n")) %>% 
  group_by(question) %>% 
  summarize(md = paste0("\n### ", question[[1]], "\n\n", paste0(md, collapse = '\n\n'))) %>% 
  ungroup() %>% 
  arrange(mixedrank(question)) %>% 
  summarize(md = paste0(md, collapse='\n')) %>% 
  pull(md) %>% clipr::write_clip()
```


