
```{r vis-setup, include=FALSE}
library(tidyverse)
```

# Visualization

We start with visualization because, well, you can see the results.

## Reading

To design good visuals, you need both *why*s and *how*s. You may have come here for the *how*s, but both are important. Our tools are changing more rapidly than ever, so if we want knowledge that lasts, we really need to know the *why*.

### Why

Read **[Look at Data](https://socviz.co/lookatdata.html)** from Healy "Data Visualization".

The text is wordy but well organized, so your speed reading skills should work well. Look at the examples: can you explain to someone else what those examples show?

### How

Read [**Data Visualization**](https://moderndive.com/2-viz.html) from ModernDive.

Try to actually answer the "Learning Check" questions for yourself. Yes this takes longer than just skimming right past them. But they may show up on a quiz...

## Application

* You did some visualization in Lab 1. How did that exercise relate to the "why" reading?

## References

* If you're the slides type, [here's some slides](https://datasciencebox.org/exploring-data.html#slides-application-exercises).
We got to most of this material but not quite all of it yet.
* <https://socviz.co/>
* [Fundamentals of Data Visualization](https://clauswilke.com/dataviz/)

## Tweaks

### Reordering bars in a bar plot

Use `fct_reorder` on the categorical variable.

```{r mass-by-species}
starwars %>% 
  drop_na(height) %>% 
  ggplot(aes(x = height, y = species)) +
  geom_boxplot()
```

```{r species-by-height-median}
starwars %>% 
  drop_na(height) %>% 
  ggplot(aes(x = height, y = fct_reorder(species, height))) +
  geom_boxplot()
```

```{r species-by-height-max}
starwars %>% 
  drop_na(height) %>% 
  ggplot(aes(x = height, y = fct_reorder(species, height, .fun = max))) +
  geom_boxplot()
```

For more info, see the [forcats vignette](https://cran.r-project.org/web/packages/forcats/vignettes/forcats.html).

### Tweaking scales

A common request: scientific notation vs not. A few options:

1. Use different units. e.g., millions of people.

```{r pop-millions-unit}
gapminder::gapminder %>% 
  filter(country == "United States") %>% 
  ggplot(aes(x = year, y = pop / 1e6)) +
  geom_line() +
  labs(y = "Population (millions)")
```

2. Use `scale_y_continuous` with `labels = scales::comma`.

```{r pop-scales-comma}
gapminder::gapminder %>% 
  filter(country == "United States") %>% 
  ggplot(aes(x = year, y = pop)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) + 
  labs(y = "Population")
```

3. Use `scales::label_number` for even more control (see the help page).

```{r pop-label-number}
gapminder::gapminder %>% 
  filter(country == "United States") %>% 
  ggplot(aes(x = year, y = pop)) +
  geom_line() +
  scale_y_continuous(labels = scales::label_number(scale = 1e-6, suffix = "M")) + 
  labs(y = "Population")
```


## Mapping

### Plotly

This document shows examples of two simple mapping tasks using Plotly. More details are available in the [plotly-r book](https://plotly-r.com/maps.html).

We'll be using the `tidyverse` and the `plotly` package.

```{r message=FALSE}
library(tidyverse)
library(plotly)
```

#### Markers

When you just want to mark something on a map, you can give lat/long coordinates to `add_markers`.

For example, let's use a dataset of US cities:

```{r}
maps::us.cities %>% head()
```

Here's how to draw it on a map.

```{r geo-markers-example}
maps::us.cities %>% 
  # Fix the column names.
  rename(state = country.etc) %>%
  # Keep only larger cities.
  filter(pop > 100000) %>% 
  # Construct the "geo" projection.
  plot_geo() %>%
  # Add state markers
  add_markers(
    # Set marker position.
    x = ~long, 
    y = ~lat, 
    # Set other aesthetics (here, redundantly encode population)
    size = ~pop, 
    color = ~pop,
    # Customize the label.
    text = ~ glue::glue("{name}, population {scales::comma(pop)}"),  
    hoverinfo = "text"
  ) %>% 
  layout(
    # Zoom into just USA.
    geo = list(
      scope = 'usa'
    )
  )
```

#### Choropleths

Plotly has builtin support for countries and US states. Any other granularity requires manually working with GeoJSON files; see
the documentation.

Let's make a world population map. First, let's construct a dataset of the most recent data that Gapminder has for each country:

```{r choropleth-example-data}
library(gapminder)
latest_country_data <- gapminder::gapminder_unfiltered %>%
  arrange(year) %>% 
  group_by(country) %>%
  slice_tail(n = 1) %>% 
  left_join(gapminder::country_codes, by = "country")
```

Now we add a "choropleth" trace. Note that this has the typical problem of choropleth maps and densities; see
[Fundamentals of Data Visualization](https://clauswilke.com/dataviz/geospatial-data.html#choropleth-mapping)
for some discussion of this.

```{r choropleth-example}
latest_country_data %>% 
  plot_geo() %>% 
  add_trace(
    type = "choropleth",
    # Specify that the "country" column contains the country names.
    locations = ~country,
    locationmode = "country names",
    # Use fill to show population. (I don't know why it's called 'z' and not 'fill'.)
    z = ~pop
  )
```



```{r include=FALSE}
#geojson <- pins::pin("https://raw.githubusercontent.com/datasets/geo-countries/master/data/countries.geojson") %>% 
#jsonlite::fromJSON(simplifyVector = FALSE)
```
