---
title: "Lab 10: Model Tuning"
output: 
  tufte::tufte_html:
    css: ../lab.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(rpart.plot)
theme_set(theme_bw())
options(scipen = 5) # encourage metrics to print in fixed-point notation
options(dplyr.summarise.inform = FALSE) # silence a warning message
```

## Goals

* Design an accurate predictive model
* Compare model performance using cross-validation
* Validate model performance on held-out data
* Practice more visualization (as always!)

## Getting started

* Pick a team to work with--ideally 2, no more than 3.
  * A team should be from within a single cohort.
  * (Crossing cohort lines is possible, but you'll need to ask the course staff to make an appropriate repo.)
* Go to [our GitHub organization](https://github.com/Calvin-DS202-FA20/).
  You'll find 5 `lab10` repos available to your cohort. Pick one for your team,
  making sure that no two teams share the same repo.
* On the attendance sheet, each team member should fill in the name of their team repo under the "Team" column. It should look like `lab10-V2`.
* Each team member should clone the repo as usual.

### Optional, experimental workflow

The RStudio server lets multiple people edit a single project at the same time!
This can be confusing since it's different from collaborating using GitHub, so
try it only if all of your team is game to try:

* The person whose name appears first on the Attendance sheet (or someone else if needed)
  clones the team's repo on the RStudio Server.
* They Share the project with the other team members
* They copy the URL and paste it into the attendance sheet
* Others go to that URL.

## Data and Initial Split

We'll again be using the Ames home sales dataset.
Again, see [Data dictionary](http://jse.amstat.org/v19n3/decock/DataDocumentation.txt).
And again, we use a subset of the original data as suggested by the
[original paper](http://jse.amstat.org/v19n3/decock.pdf).

We'll also scale the sale price to be in units of $1000, to keep the numbers more manageable.

```{r load-and-subset-data}
data(ames, package = "modeldata")
ames_all <- ames %>%
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal") %>%
  mutate(across(where(is.integer), as.double)) %>%
  mutate(Sale_Price = Sale_Price / 1000)
rm(ames)
```

We'll tweak the train-test split to give us a slightly larger training set:

```{r train-test-split}
set.seed(10)
ames_split <- initial_split(ames_all, prop = 3 / 4)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

## Modeling

Now that we have techniques for specifying a model, preprocessing data to give to it,
and estimating its performance, let's try to just make a good model.

We learned about *recipes* and *workflows* but only used them occasonially last week.
For this lab, though, they will become essential as we seek to compare models
that use different variables, different preprocessing steps, different models,
etc, using the same methodology.

Our basic strategy will be to collect model specs and metrics in a data frame.
We'll run through the approach once for a basic linreg and decision tree,
then you'll try out things on your own.

### Initialize

Overall let's use 10-fold cross-validation.

```{r declare-cv}
ames_resamples <- ames_train %>% 
  vfold_cv(v = 10)
```

These models can be time-consuming to run, so let's collect the results in a
`modeling_results` data frame as we go. Let's start off with no results:

```{r empty-modeling-results}
modeling_results <- tibble()
```

And here's a function to add a new set of results. Don't worry about how it works.

```{r append-new-results}
append_new_results <- function(modeling_results, model_name, spec, cv_results) {
  if (nrow(modeling_results) > 0 && model_name %in% modeling_results$model_name)
    stop("Model with that name already exists.")
  
  bind_rows(
    modeling_results %>% mutate(model_name = as.character(model_name)),
    tibble(
      model_name = model_name,
      spec = list(spec),
      cv_results %>% select(-.estimator)
    )
  ) %>% 
    mutate(model_name = as_factor(model_name)) # Ensure that factor level matches insertion order.
}

get_spec_for_model <- function(model_name) {
  modeling_results %>% filter(model_name == !!model_name) %>% purrr::chuck("spec", 1)
}
```

Finally, a few utilities:

```{r utilities}
add_predictions <- function(data, ...) {
  imap_dfr(
    rlang::dots_list(..., .named = TRUE),
    function(model, model_name) {
      model %>%
        predict(data) %>%
        bind_cols(data) %>%
        mutate(model = !!model_name)
    }
  )
}

sweep_model_examples <- function(model, dataset, vars_to_sweep, examples = slice_sample(dataset, n = 10)) {
  X <- map_dfr(vars_to_sweep, function(v) {
    var_to_sweep <- rlang::as_label(v)
    sweep_min <- min(dataset[[var_to_sweep]])
    sweep_max <- max(dataset[[var_to_sweep]])
    expand_grid(
      examples %>% select(-!!var_to_sweep) %>% mutate(.idx = row_number()),
      !!enquo(var_to_sweep) := seq(sweep_min, sweep_max, length.out = 500)) %>% 
      mutate(sweep_var = var_to_sweep, .sweep_val = .data[[var_to_sweep]])
  })
  model %>% 
    predict(X) %>% 
    bind_cols(X)
}

linear_reg <- function(engine = "lm", ...) {
  parsnip::linear_reg(...) %>% set_engine(engine)
}

decision_tree <- function(mode = "regression", engine = "rpart", ...) {
  parsnip::decision_tree(mode = "regression", ...) %>%
    set_engine(engine)
}
```


### Linear Regression

Let's first fit a simple linear regression. First we define our spec.

```{marginfigure}
Note that we need to give `recipe` a `data` parameter. This is the *template*
data, just used to tell the recipe what columns there are and what data type
each one has. It does not actually train on this data!
```


```{r linreg1-spec}
spec <- workflow() %>% 
  add_recipe(
    recipe(Sale_Price ~ Latitude + Longitude, data = ames_train)
  ) %>% 
  add_model(
    linear_reg()
  )
```

Then, let's get the cross-validation assessment scores:

```{r linreg1-cv}
cv_results <- 
  spec %>% 
  fit_resamples(resamples = ames_resamples, metrics = metric_set(mae)) %>% 
  collect_metrics(summarize = FALSE)

cv_results
```

Finally, we collect these results into our `modeling_results` data frame.

```{r linreg1-collect}
modeling_results <- modeling_results %>% 
  append_new_results(
    model_name = "Simple Linreg",
    spec = spec,
    cv_results = cv_results
  )
modeling_results
```

### Decision Tree

Your turn: repeat the above, but now using the default `decision_tree()`.

```{r tree1-spec, echo=FALSE}
spec <- workflow() %>% 
  add_recipe(
    recipe(Sale_Price ~ Latitude + Longitude, data = ames_train)
  ) %>% 
  add_model(
    decision_tree()
  )
```

```{r tree1-cv, echo=FALSE}
cv_results <- 
  spec %>% 
  fit_resamples(resamples = ames_resamples, metrics = metric_set(mae)) %>% 
  collect_metrics(summarize = FALSE)
```

```{r tree1-collect, echo=FALSE}
modeling_results <- modeling_results %>% 
  append_new_results(
    model_name = "Decision Tree",
    spec = spec,
    cv_results = cv_results
  )
modeling_results
```

## Compare model performance

Compare the performance of the two models. Which one is better?

```{r compare-models}
modeling_results %>% 
  ggplot(aes(y = fct_rev(model_name), x = .estimate)) +
    geom_boxplot() +
    geom_point() +
    labs(x = "Mean Absolute Error ($1000)", y = "")
```


### Better Decision Tree

Try getting a better decision tree model:

* Use more features
* Tweak the `decision_tree` parameters (run `?rpart::rpart.control` for details)
  * `cost_complexity`: any split that does not improve the fit by this much is not attempted (defaults to 0.01, aka `cp`)
  * `tree_depth`: maximum depth of any node in the final tree (defaults to 30, aka `maxdepth`)
  * `min_n`: minimum number of observations that must exist in a node in order for a split to be attempted (defaults to 20, aka `minsplit`)

```{r tree2-spec, echo=FALSE}
spec <- workflow() %>% 
  add_recipe(
    recipe(Sale_Price ~ Latitude + Longitude + Gr_Liv_Area + Bldg_Type, data = ames_train)
  ) %>% 
  add_model(
    decision_tree()
  )
```

```{r tree2-cv, echo=FALSE}
cv_results <- 
  spec %>% 
  fit_resamples(resamples = ames_resamples, metrics = metric_set(mae)) %>% 
  collect_metrics(summarize = FALSE)
```

```{r tree2-collect, echo=FALSE}
modeling_results <- modeling_results %>% 
  append_new_results(
    model_name = "Decision Tree 2",#: Lat, Long, LivArea, BldgType",
    spec = spec,
    cv_results = cv_results
  )
```

## Visualize models

Let's look at two ways of visualizing what's going on in fitted models.
The examples in this section will show the results of the last model that
your instructor happened to spec out (while playing around with decision trees
for the last section); notice what's different when you apply these
techniques to different models.

First, let's get a model fit on the whole training set. Remember: the cross validation process trained a bunch of smaller models; now we want to look at the slightly bigger full model.

```{r fit-on-full-train}
model <- 
  get_spec_for_model("Decision Tree 2") %>%
  fit(ames_train)
```

#### Examine Residuals

Here's a way to see at a glance what might be helpful features to add becasue
there are patterns in the residuals. We just plot the residuals against all
of the other numeric features. (It doesn't handle categorical features, but 
you can try mapping a categorical feature to color or something.)

```{r resids-against-all-numeric, fig.width=10, fig.height=10, fig.fullwidth=TRUE}
ames_train %>% 
  add_predictions(model) %>% 
  mutate(resid = Sale_Price - .pred) %>% 
  # Pivot all of the numeric variables except for the outcomes and residuals.
  pivot_longer(c(where(is.numeric) & !any_of(c("Sale_Price", ".pred", "resid")))) %>% 
  # Plot!
  ggplot(aes(x = value, y = resid)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(vars(name), scales = "free")
```

##### Examine Predictions

The `sweep_model_examples` function defined above creates example predictions
by starting with 10 randomly selected example data points and sweeping a variable
(columns) across the whole range of values seen in the training set.
You give it one or more variables using `vars`, the same quoting
function that `facet_wrap` uses. **Note**: it currently only works for continuous
variables.

The following example shows how to sweep several variables and plot each sweep
as a line.

```{r sweep-model-examples, fig.fullwidth=TRUE}
set.seed(0)
sweep_model_examples(model, ames_train, vars(Longitude, Latitude, Gr_Liv_Area)) %>% 
  ggplot(aes(x = .sweep_val, y = .pred, group = .idx, color = Latitude)) +
  geom_line(alpha = .5) +
  facet_wrap(vars(sweep_var), scales = "free")
```



### Better Linreg

Now, try to get a better linear regression model. You might try:

* Use more features
  * add dummy encoding steps as needed (`step_dummy`, and perhaps `step_other` if needed)
* Transform existing features
  * `step_poly`, or its better-behaved cousin `step_ns`
  * `step_discretize`
  * `step_interact` (see hw9 for an example)

```{r linreg2-spec, echo=FALSE}
spec <- workflow() %>% 
  add_recipe(
    recipe(Sale_Price ~ Latitude + Longitude + Gr_Liv_Area + Bldg_Type, data = ames_train) %>% 
      step_other(Bldg_Type) %>% 
      step_dummy(Bldg_Type) %>% 
      step_ns(Latitude) %>% 
      step_ns(Longitude)
  ) %>% 
  add_model(
    linear_reg()
  )
```

```{r linreg2-cv, echo=FALSE}
cv_results <- 
  spec %>% 
  fit_resamples(resamples = ames_resamples, metrics = metric_set(mae)) %>% 
  collect_metrics(summarize = FALSE)
```

```{r linreg2-collect, echo=FALSE}
modeling_results <- modeling_results %>% 
  append_new_results(
    model_name = "LinReg 2",#: Lat, Long, LivArea, BldgType, natsplines on LatLong",
    spec = spec,
    cv_results = cv_results
  )
```



## Compare model performance

The other visuals are just for you to explore. Keep your model performance visual 
chunk after all of your modeling results have been collected, so that you can
plot the performance here.

For example, here is the result with some models I was playing around with.

```{r}
modeling_results %>% 
  ggplot(aes(y = fct_rev(model_name), x = .estimate)) +
    geom_boxplot() +
    geom_point() +
    labs(x = "Mean Absolute Error ($1000)", y = "")
```

