---
title: "HW 1"
output: 
  tufte::tufte_html:
    css: ../hw.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggformula)
```

## Learning Goals

## Getting Help

If you have any questions about the assignment, please post them on Teams!

## Getting started

Click [HERE](https://classroom.github.com/a/eLXKhluX) to claim your repo for this assignment.

First, open the R Markdown document `hw-01-welcome.Rmd` and Knit it. Make sure it compiles without errors. The output will be in the `.md` file with the same name.

The document contains placeholders for the following exercises.

```{r wrangle, include=FALSE}
daily_rides <- read_csv("../../data/bikeshare/day.csv") %>%
  mutate(workingday = as.factor(case_when(
    workingday == 1 ~ "workday",
    TRUE ~ "weekend"))) %>% 
  select(
    date = dteday,
    day_of_week = weekday,
    workingday,
    temp,
    casual, registered, total = cnt
  )
#write_csv(daily_rides, "../../data/bikeshare/day_simple.csv")
daily_rides %>% glimpse()
```

# The Data

Let's imagine we're hired by the administrators of the [Capital Bikeshare program](https://www.capitalbikeshare.com) program in Washington D.C., to **help them predict the hourly demand for rental bikes**.. We will be working with this data over the next several homework assignments.

The predicted hourly demand could be used for planning the number of bikes that need to be available in the system at any given hour of the day. It costs the program money if bike stations are full and bikes cannot be returned, or empty and there are no bikes available. You will use multiple linear regression and polynomial regression and will explore techniques for subset selection to predict bike usage. Our eventual goal will be to build a regression model that can predict the total number of bike rentals in a given hour of the day, based on all the information we can get.

The data for this problem were collected from the Capital Bikeshare program over the course of two years (2011 and 2012). Researchers at the University of Porto processed the data and augmented it with extra information, as described [on this page](https://archive.ics.uci.edu/ml/datasets/bike+sharing+dataset). It has been cleaned up slightly for this assignment.


## Load and Explore Data

The data is in a CSV (comma separated values) file called `day.csv` in the `data/` folder in your repository. You can load this file into R using the `read_csv()` function. The following code block is already included.

```{r load-data, eval=FALSE, message=FALSE}
daily_rides <- read_csv("data/day_simple.csv")
```

Practice reading that out loud: "`daily_rides` gets `read_csv` of `data/day_simple.csv`." (That's a *forward* slash.)

1. Take a quick look at the data: you'll write R code that will be read "take `daily_rides` and then `glimpse()` it."

The columns are:

* `date`
* `day_of_week`: an integer between 0 and 6 inclusive. In a later exercise, you will decode which number represents Monday, etc.
* `workingday`: either "weekday" or "weekend" (which actually includes holidays too).
* `temp`: the average temperature that day, scaled so that the lowest temperature in the dataset is 0 and the highest is 1.
* `registered`: the number of rides by users who registered for a Capital Bikeshare account.
* `casual`: the number of rides by "casual" users (who didn't register for a Capital Bikeshare account)
* `total`: `registered` + `casual`.

2. How many rows does the dataset contain? What does each row represent?

3. Fill in the blanks provided to create the following plot, a scatter plot showing the number of rides by *registered* users each day.
(Try removing `geom_point()` and `geom_smooth()` one at a time to make sure you understand what each one does.)

```{r registered-by-day, echo=FALSE, message=FALSE}
daily_rides %>% 
  ggplot(aes(x = date, y = registered)) +
    geom_point() +
    geom_smooth() +
    labs(x = "date", y = "number of rides by registered users")
```

`r emo::ji("white_check_mark")` `r emo::ji("arrow_up")` *Now is a good time to commit and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.*

4. Make the same plot again here, but add a mapping of `workingday` to the `color` aesthetic. Your result should look like:

```{r registered-by-day-workingday, echo=FALSE, message=FALSE}
daily_rides %>% 
  ggplot(aes(x = date, y = registered, color = workingday)) +
    geom_point() +
    geom_smooth() +
    labs(x = "date", y = "number of rides by registered users", color = "")
```

5. Write a one or two sentence interpretation of the graph, focusing on the following question: *How do the number of rides by registered users compare for weekdays vs weekends?*

6. Make a scatter plot showing the how the number of rides by registered users varies with temperature.
Design the plot so that it shows the difference between weekends and weekdays.

```{r registered-by-temp, echo=FALSE, message=FALSE}
daily_rides %>% 
  ggplot(aes(x = temp, y = registered, color = workingday)) +
    geom_point() +
    geom_smooth() +
    labs(x = "normalized temperature", y = "number of rides by registered users", color="")

```


