---
title: "HW 9: Predictive Modeling"
output: 
  tufte::tufte_html:
    css: ../hw.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidymodels)
theme_set(theme_bw())
options(dplyr.summarise.inform = FALSE) # silence a warning message
```

## Goals

* Apply predictive modeling using `tidymodels` to a time-series prediction task
* Use visualization to understand the strengths and weaknesses of different models
* Compare and contrast linear regression and decision trees

Here are some useful utility functions. They will let you omit the `set_engine`
step; just use `linear_reg() %>% fit(...)`.

```{r echo=TRUE}
add_predictions <- function(data, ...) {
  model_names <- as.character(eval(substitute(alist(...))))
  models <- list(...)
  map2_dfr(
    model_names,
    models,
    function(model_name, model) {
      model %>%
        predict(data) %>% 
        bind_cols(data) %>% 
        mutate(model = !!model_name)
    }
  )
}

linear_reg <- function(engine = "lm", ...) { 
  parsnip::linear_reg(...) %>% set_engine(engine)
}
decision_tree <- function(mode = "regression", engine = "rpart", ...) {
  parsnip::decision_tree(mode = "regression", ...) %>%
    set_engine(engine)
}
```

## Getting Started

### Load data

```{r echo=TRUE}
daily_rides <- read_csv("../../data/bikeshare/day.csv", col_types = cols(
  .default = col_double(),
  dteday = col_date()
)) %>% 
  mutate(
    yr = as_factor(2011 + yr),
    weekday = lubridate::wday(dteday, label = TRUE),
    workingday = as_factor(if_else(workingday == 1, "workday", "weekend")),
    temp = temp * (39 + 8) - 8,
    atemp = atemp * (50 + 16) - 16
    ) %>% 
  rename(
    year = yr,
    date = dteday
  )
```

### Exploratory Analytics

Make a scatter plot of the number of `casual` riders by `date`, color-coding
whether each day is a `workingday` or not.

```{r casual-vs-day-eda}
daily_rides %>%
  ggplot(aes(x = date, y = casual, color = workingday)) +
  geom_point()
```

### Train-test split

```{r}
train <- daily_rides %>% filter(year == 2011)
test <- daily_rides %>% filter(year == 2012)
```

Check that you get `r nrow(train)` training set days and `r nrow(test)` test set days. (why might this make sense?)

## Linear Regression using Temperature

Fit a linear regression model to predict the number of `casual` riders from the `temp`erature.
Use the training dataset (2011). Call the model `model1`.

```{r}
model1 <- linear_reg() %>%
  fit(casual ~ temp, data = train)
```

### Look inside the model

We can use the `tidy` function to get the model's coefficients:

```{r echo=TRUE}
model1 %>% tidy() %>% select(term, estimate)
```

Complete this sentence: "For every additional degree C, model1 predicts ___ additional riders."

### Predictions

Plot the model's predictions along with the number of casual rides.

(Start by piping `train` into `add_predictions(model1)`. You'll want to add a `geom_line`
where you specify `.pred` as the `y` aesthetic. This works because the dataset is sorted by `date` already.)

```{r show-model1}
train %>%
  add_predictions(model1) %>%
  ggplot(aes(x = date, y = casual, color = workingday)) +
  geom_point() +
  geom_line(aes(y = .pred), color = "maroon", size = 1)
```

This is a linear model, so generally we expect to see straight lines. Why is
the maroon line not a straight line?

### Residuals Histogram

Make a histogram of the residuals, faceted by weekend vs weekday. (The example
uses a `binwidth` of 100 and a `geom_vline` at 0.)

```{r resid-histogram-model1}
train %>% 
  add_predictions(model1) %>%
  mutate(resid = casual - .pred) %>%
  ggplot(aes(x = resid)) +
  geom_histogram(binwidth = 100) +
  geom_vline(xintercept = 0) +
  facet_wrap(vars(workingday), nrow = 2, scales = "free_y")
```

### Observed by Predicted

Make an observed-by-predicted plot. Add `coord_obs_pred()` to make the axes
exactly equal, and `geom_abline` to draw the line corresponding to exactly
correct predictions.

```{r observed-by-predicted-model1}
train %>%
  add_predictions(model1) %>%
  ggplot(aes(x = casual, y = .pred, color = workingday)) +
  geom_abline() +
  geom_point(alpha = .5) +
  coord_obs_pred()
```

Under what circumstances does this this model typically predict too high? Under
what circumstances does it predict too low? Use the residual plot
to justify your answer. 

Try to guess why might the model have done that.

### Validate the model on the test set

Now, make the same plots as above, but with the full `daily_rides` dataset
(so now we have 2012 also.)

#### Predictions

```{r show-model1-test}
daily_rides %>%
  add_predictions(model1) %>%
  ggplot(aes(x = date, y = casual, color = workingday)) +
  geom_point() +
  geom_line(aes(y = .pred), color = "maroon", size = 1)
```

#### Residuals

For this plot, facet on `year` instead of `workingday`, to compare the training set and test set.

```{r}
daily_rides %>%
  add_predictions(model1) %>%
  mutate(resid = casual - .pred) %>%
  ggplot(aes(x = resid)) +
  geom_histogram(binwidth = 100) +
  facet_wrap(vars(year), nrow = 2)
```

#### Observed by Predicted

Again, color by `year` instead of `workingday`.

```{r}
daily_rides %>%
  add_predictions(model1) %>%
  ggplot(aes(x = casual, y = .pred, color = year)) +
  geom_abline() +
  geom_point(alpha = .5) +
  coord_obs_pred()
```

### Quantify errors

Use the `mae` function, or write your own `summary` call, to compute the Mean
Absolute Error for each year (2011 and 2012).

```{r}
daily_rides %>% 
  add_predictions(model1) %>% 
  group_by(year) %>% 
  summarize(mae = mean(abs(casual - .pred)))
#  mae(truth = casual, estimate = .pred)
```

How incorrect was the prediction, on average, for each year?

#### Summarize

Write one-sentence responses for the following:

* How well does this model perform on the training set?
* How does its test set performance compare with its training set performance?
* Compare the "quantify errors" table above with the plots of model performance.
  * What can you easily observe from the plots but not the table?
  * What can you easily observe from the table but not the plots?

## Predict from temperature and day of week

Since it seems that `workingday` makes a big difference in the number of casual
riders, let's include it in our model. It's a categorical variable, so we'll
need to dummy-encode it.

Dummy encoding will only give us a conditional *intercept*. If we want our 
*slope* to be conditional also, we need to make an interaction term.

```{r echo=TRUE}
recipe2 <- 
  recipe(casual ~ temp + workingday, data = train) %>% 
  step_dummy(workingday) %>% 
  step_interact(~ temp:starts_with("workingday"))
  

model2 <- workflow() %>% 
  add_recipe(recipe2) %>% 
  add_model(linear_reg()) %>% 
  fit(train)
model2
```

```{r}
train %>%
  add_predictions(model2) %>%
  ggplot(aes(x = date, y = casual, color = workingday)) +
  geom_point() +
  geom_line(aes(y = .pred), size = 1)
```


```{r}
train %>% 
  ggplot(aes(x = temp, y = casual, color = workingday)) +
    geom_point() +
    geom_line(
      data = expand_grid(
        workingday = levels(train$workingday),
        temp = modelr::seq_range(train$temp, n = 100)
      ) %>% 
        add_predictions(model2),
      mapping = aes(y = .pred)
    )
```



```{r}
daily_rides %>% 
  add_predictions(model2) %>% 
  group_by(year) %>% 
  mae(truth = casual, estimate = .pred)
```


## Try decision tree

```{r}
model3 <- 
  decision_tree() %>% 
  fit(casual ~ temp + workingday, data = train)
model3
```

```{r}
train %>%
  add_predictions(model3) %>%
  ggplot(aes(x = date, y = casual, color = workingday)) +
  geom_point() +
  geom_line(aes(y = .pred), size = 1)
```


```{r}
daily_rides %>% 
  add_predictions(model3) %>% 
  group_by(year) %>% 
  mae(truth = casual, estimate = .pred)
```

```{r}
daily_rides %>% 
  add_predictions(model1, model2, model3) %>% 
  group_by(model, year) %>% 
  summarize(mae = mean(abs(casual - .pred))) %>% 
  ggplot(aes(x = model, y = mae, fill = year)) +
    geom_col(position = 'dodge')

```

