---
title: "HW 9: Predictive Modeling"
output: 
  tufte::tufte_html:
    css: ../hw.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidymodels)
```

## Goals

* Apply predictive modeling using `tidymodels` to a time-series prediction task
* Compare and contrast linear regression and decision trees

Here are some useful utility functions. They will let you omit the `set_engine`
step; just use `linear_reg() %>% fit(...)`.

```{r echo=TRUE}
add_predictions <- function(data, model, variable_name = ".pred", model_name = deparse(substitute(model))) {
  model %>%
    predict(data) %>%
    rename(!!enquo(variable_name) := .pred) %>%
    mutate(model = model_name) %>%
    bind_cols(data)
}

linear_reg <- function(engine = "lm", ...) { 
  parsnip::linear_reg(...) %>% set_engine(engine)
}
decision_tree <- function(mode = "regression", engine = "rpart", ...) {
  parsnip::decision_tree(mode = "regression", ...) %>%
    set_engine(engine)
}
```

## Load data

```{r echo=TRUE}
daily_rides <- read_csv("../../data/bikeshare/day.csv", col_types = cols(
  .default = col_double(),
  dteday = col_date()
)) %>% 
  mutate(
    yr = as_factor(2011 + yr),
    weekday = lubridate::wday(dteday, label = TRUE),
    workingday = as_factor(workingday),
    temp = temp * (39 + 8) - 8,
    atemp = atemp * (50 + 16) - 16
    ) %>% 
  rename(
    year = yr,
    date = dteday
  )
```

## Exploratory Analytics

Make a scatterplot of the number of `casual` riders by `date`, color-coding
whether each day is a `workingday` or not.

```{r casual-vs-day-eda}
daily_rides %>%
  ggplot(aes(x = date, y = casual, color = workingday)) +
  geom_point()
```

### Train-test split

```{r}
train <- daily_rides %>% filter(year == 2011)
test <- daily_rides %>% filter(year == 2012)
```

Check that you get `r nrow(train)` training set days and `r nrow(test)` test set days. (why might this make sense?)

## Predict from temperature

Fit a linear regression model to predict the number of `casual` riders from the `temp`erature.
Use the training dataset (2011). Call the model `model1`.

```{r}
model1 <- linear_reg() %>%
  fit(casual ~ temp, data = train)
```

We can use the `tidy` function to get the model's coefficients:

```{r echo=TRUE}
model1 %>% tidy() %>% select(term, estimate)
```

Complete this sentence: "For every additional degree C, model1 predicts ___ additional riders."

Plot the model's predictions along with the number of casual rides.

(Start by piping `train` into `add_predictions(model1)`. You'll want to add a `geom_line`
where you specify `.pred` as the `y` aesthetic. This works because the dataset is sorted by `date` already.)

```{r show-model1}
train %>%
  add_predictions(model1) %>%
  ggplot(aes(x = date, y = casual, color = workingday)) +
  geom_point() +
  geom_line(aes(y = .pred), color = "maroon", size = 1)
```

This is a linear model, so generally we expect to see straight lines. Why is
the maroon line not a straight line?

### Look at residuals

```{r}
train %>% 
  add_predictions(model1) %>%
  mutate(resid = casual - .pred) %>%
  ggplot(aes(x = resid)) +
  geom_histogram(binwidth = 100) +
  facet_wrap(vars(workingday), nrow = 2)
```



```{r}
train %>%
  add_predictions(model1) %>%
  ggplot(aes(x = casual, y = .pred, color = workingday)) +
  geom_abline() +
  geom_point(alpha = .5) +
  coord_obs_pred()
```

Under what circumstances does this this model typically predict too high? Under
what circumstances does it predict too low? Use the residual plot
to justify your answer. 

Why might the model have done that?


## Validate on 2012

```{r}
daily_rides %>%
  add_predictions(model1) %>%
  ggplot(aes(x = casual, y = .pred, color = year)) +
  geom_abline() +
  geom_point(alpha = .5) +
  coord_obs_pred()
```

```{r}
daily_rides %>%
  add_predictions(model1) %>%
  mutate(resid = casual - .pred) %>%
  ggplot(aes(x = resid)) +
  geom_histogram(binwidth = 100) +
  facet_wrap(vars(year), nrow = 2)
```

```{r}
daily_rides %>% 
  add_predictions(model1) %>% 
  group_by(year) %>% 
  mae(truth = casual, estimate = .pred)
```

How incorrect was the prediction, on average, for each year?

How does looking at a single number compare with looking at the plots?

## Predict from temperature and day of week

```{r}
recipe2 <- 
  recipe(casual ~ temp + workingday, data = train) %>% 
  step_dummy(workingday) %>% 
  step_interact(~ temp:starts_with("workingday"))
  

model2 <- workflow() %>% 
  add_recipe(recipe2) %>% 
  add_model(linear_reg()) %>% 
  fit(train)
model2
```

```{r}
train %>%
  add_predictions(model2) %>%
  ggplot(aes(x = date, y = casual, color = workingday)) +
  geom_point() +
  geom_line(aes(y = .pred), size = 1)
```


```{r}
train %>% 
  ggplot(aes(x = temp, y = casual, color = workingday)) +
    geom_point() +
    geom_line(
      data = expand_grid(
        workingday = levels(train$workingday),
        temp = modelr::seq_range(train$temp, n = 100)
      ) %>% 
        add_predictions(model2),
      mapping = aes(y = .pred)
    )
```



```{r}
daily_rides %>% 
  add_predictions(model2) %>% 
  group_by(year) %>% 
  mae(truth = casual, estimate = .pred)
```


## Try decision tree

```{r}
model3 <- 
  decision_tree() %>% 
  fit(casual ~ temp + workingday, data = train)
model3
```

```{r}
train %>%
  add_predictions(model3) %>%
  ggplot(aes(x = date, y = casual, color = workingday)) +
  geom_point() +
  geom_line(aes(y = .pred), size = 1)
```


```{r}
daily_rides %>% 
  add_predictions(model3) %>% 
  group_by(year) %>% 
  mae(truth = casual, estimate = .pred)
```
