---
title: "HW 2: More Visualization Practice"
output: 
  tufte::tufte_html:
    css: ../hw.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
options(dplyr.summarise.inform = FALSE)
```

We will be continuing our work with the Capital Bikeshare dataset. Our goal
continues to be to understand the pattern of ridership better to evaluate the
current system and suggest potential improvements.

Towards that end, we will construct some more visualizations, this time using
the more fine-grained *hourly* ridership data.

## Learning goals

Upon completing this homework, you will have greater comfort making and refining
visualizations using the Grammar of Graphics. You will also gain some initial
exposure to data wrangling, which we will study in depth next week.

## Collaboration

This homework may be done individually or in pairs.

If done in pairs, each student in the pair should have at least one substantive
commit to the project repository.

## Getting help

As always, if you need help, post in Teams or come to office hours!

See also the "Data Visualization with ggplot2" cheatsheet available under the RStudio Help menu.

## Getting started

**Click to [Claim your repo.](https://classroom.github.com/a/DQbjzvvK)**

You will notice that the report starts with a code block that reads the data and sets the
correct data types for several of the columns. You do not need to understand how this code works at the moment.

```{r read-data, message=FALSE, include=FALSE}
hourly_rides <- read_csv("../../data/bikeshare/hour.csv") %>%
  mutate(workingday = case_when(workingday == 1 ~ "workday",
                                          TRUE ~ "weekend"),
         ) %>%
  select(
    date = dteday,
    hour = hr,
    season,
    year = yr,
    holiday,
    workingday = workingday,
    day_of_week = weekday,
    weather_type = weathersit,
    temp,
    feels_like = atemp,
    humidity = hum,
    windspeed,
    casual,
    registered
  ) %>%
  pivot_longer(c(casual, registered),
               names_to = "member_type",
               values_to = "rides") %>% 
  mutate(across(c(season, year, hour, holiday, workingday, day_of_week, weather_type, member_type), as.factor))

hourly_rides %>% write_csv("../../../src/hw/hw02-vis/data/hour.csv")

# mutate(day_of_week = factor(day_of_week, levels = c(0, 1, 2, 3, 4, 5, 6), labels = c("Sun", "Mon", "Tue", "Wed", "Thur", "Fri", "Sat")))
#hourly_rides %>% glimpse()
```


# Exercise 1: Label days of week

The data set uses the integers 0 through 6 to label days of the week. It does not
document, however, what 0 means, or what 6 means. If we want to make understandable
plots, we should label these days of the week.

Fill in the blank in the given code block to give labels to the weekdays.
Include a brief description of how you concluded that those were the correct labels.

Use abbreviations ("Wed", "Fri", ...).

Hint: look at the data. (`View()`, `head()`, ... but leave a small, coherent set of evidence for your conclusion.)

```{r label-weekdays, include=FALSE}
hourly_rides <- hourly_rides %>%
  mutate(day_of_week = factor(day_of_week, levels = c(0, 1, 2, 3, 4, 5, 6), labels = c("Sun", "Mon", "Tue", "Wed", "Thur", "Fri", "Sat")))
```


# Exercise 2: Overall counts by weekday

The following code (also in your template) totals the number of rides by each
member by each day of week. We will be learning how to write code like this
next week:

```{r ride-totals}
ride_totals <- hourly_rides %>%
  group_by(member_type, day_of_week) %>% 
  summarize(rides = sum(rides))
```

Make the following plot. (You may choose to refine it, but a plot like this is acceptable.)

```{r ride-totals-by-weekday, echo=FALSE}
ggplot(ride_totals, aes(x = day_of_week, fill = member_type, y = rides)) +
  geom_col(position = position_dodge()) +
  labs(y = "Rides", x = "Day of Week", fill = "Type of Rider")
```

Hints:

* Use `geom_col`.
* To put the bars side-by-side instead of stacked, use `position = position_dodge()`.


# Exercise 3: Distribution of Rides by Weekday

The bar plot shows a single number for each weekday. Refine the graph to show
the *distribution* of rides each hour.

To do this:

* Use the original `hourly_rides` dataset.
* Use a box plot

Your plot might look something like this:

```{r ride-distributions-by-weekday, echo=FALSE}
hourly_rides %>%
  ggplot(aes(x = day_of_week, y = rides, fill = member_type)) +
  geom_boxplot() +
  labs(y = "Rides", x = "Day of Week", fill = "Type of Rider")
```

# Exercise 4: Distribution of Rides by Hour

Instead of variation by day of week, plot the variation by hour, distinguishing the type of rider and between working days and weekends. Details:

* The x axis should be hour. (Feel free to try flipping the axes too.)
* Use side-by-side box plots for casual and registered members.
* Label the plot.
* Use `facet_wrap` to make separate graphs for working days and weekends. Try different settings for the scales and number of rows / columns until you find a layout that looks good. (Read the help page for `facet_wrap` to find out what arguments to use.)
* Set the `fig.height` and/or `fig.width` cell options to make a comfortably sized graph.

```{r ride-distributions-by-hour, fig.height=10, fig.width=10, eval=FALSE, include=FALSE}
ggplot(hourly_rides, aes(x = hour, y = rides, fill = member_type)) + 
  geom_boxplot() +
  facet_wrap(~ workingday, scales = "free_y", nrow = 2)
```

# Exercise 5: Overplotting

The following plot attempts to show how temperature affects ridership. However, it suffers
from *overplotting*.

```{r rides-by-temp-overplot}
ggplot(hourly_rides, aes(x = temp, y = rides)) +
  geom_point() +
  facet_wrap(~ member_type) + 
  labs(x = "Normalized temperature", y = "Rides")
```

Pick at least two of the following methods to try to fix the overplotting issue. Which do you prefer, and why? Include the code and results of both plots you tried.

* `geom_density_2d`
* Making points transparent (`alpha = 0.05`) or very small (`shape = '.'`)
* binning the x axis (`group = cut_width(temp, ___)`)
* any other approach you want to try!

Note: you don't need to make proper labels for these plots, though it's always nice.

```{r binned-x, include=FALSE}
hourly_rides %>%
  ggplot(aes(x = temp, y = rides, group = cut_width(temp, .1))) +
    geom_boxplot() +
    facet_wrap(~ member_type)
```
