---
title: "HW 2: More Visualization Practice"
output: 
  tufte::tufte_html:
    css: ../hw.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
options(dplyr.summarise.inform = FALSE)
```

We will be continuing our work with the Capital Bikeshare dataset. Our goal
continues to be to understand the pattern of ridership better to evaluate the
current system and suggest potential improvements.

Towards that end, we will construct some more visualizations, this time using
the more fine-grained *hourly* ridership data.

## Learning goals

Getting proficient at designing and implementing data visualizations takes practice.
The goal of this assignment is to give you practice specifically with:

* Analyzing a visualization in terms of the Grammar of Graphics
* Constructing a visualization using `ggplot`
* Choosing the the structure of a visualization according to characteristics of the data and questions to be answered.

You will also gain practice at:

* Making data-driven arguments (Exercise 1)

Upon completing this homework, you will have greater comfort making and refining
visualizations using the Grammar of Graphics. You will also gain some initial
exposure to data wrangling, which we will study in depth next week.

## Tips for Success

* Work at this a little bit at a time.
* If you're stuck on some coding aspect, ask for help soon: don't spend more than 15 minutes trying to figure out a
  coding problem.
* As always, if you need help, post in Piazza or come to office hours!

See also the "Data Visualization with ggplot2" cheatsheet available under the RStudio Help menu and
other resources linked on the Materials section of the course Syllabus.

## Getting started

Pull your Portfolio repo and find `hw/hw02/hw02-vis.Rmd`.

Update the title and author, Knit, Commit, Push.

### Exercise 0: Getting to know the data

You will notice that the report starts with a code block that reads the data and sets the
correct data types for several of the columns. You do not need to understand how this code works at the moment.

```{r read-data, message=FALSE, include=FALSE}
roundtrip_csv <- function(data, filename) {
  write_csv(data, filename)
  read_csv(filename)
}

daily_rides <- read_csv(here::here("docs/data/bikeshare/day.csv")) %>%
  mutate(workingday = case_when(workingday == 1 ~ "workday",
                                          TRUE ~ "weekend"),
         ) %>%
  select(
    date = dteday,
    season,
    year = yr,
    holiday,
    workingday = workingday,
    day_of_week = weekday,
    weather_type = weathersit,
    temp,
    feels_like = atemp,
    humidity = hum,
    windspeed,
    casual,
    registered
  ) %>%
  mutate(season = c("W", "S", "U", "F")[season]) %>% 
  pivot_longer(c(casual, registered),
               names_to = "rider_type",
               values_to = "rides") %>% 
  relocate(date, rider_type, rides) %>% 
  mutate(across(c(season, year, holiday, workingday, day_of_week, weather_type, rider_type), as.factor))

daily_rides <- roundtrip_csv(daily_rides, here::here("src/hw/hw02-vis/data/day.csv"))

if(FALSE) {
hourly_rides <- read_csv("../../data/bikeshare/hour.csv") %>%
  mutate(workingday = case_when(workingday == 1 ~ "workday",
                                          TRUE ~ "weekend"),
         ) %>%
  select(
    date = dteday,
    hour = hr,
    season,
    year = yr,
    holiday,
    workingday = workingday,
    day_of_week = weekday,
    weather_type = weathersit,
    temp,
    feels_like = atemp,
    humidity = hum,
    windspeed,
    casual,
    registered
  ) %>%
  mutate(season = c("W", "S", "U", "F")[season]) %>% 
  pivot_longer(c(casual, registered),
               names_to = "rider_type",
               values_to = "rides") %>% 
  relocate(date, hour, rider_type, rides) %>% 
  mutate(across(c(season, year, hour, holiday, workingday, day_of_week, weather_type, rider_type), as.factor))

hourly_rides %>% write_csv(here::here("src/hw/hw02-vis/data/hour.csv"))
hourly_rides <- read_csv(here::here("src/hw/hw02-vis/data/hour.csv"))
}
```

Notice also that the template shows two different ways shown of getting a quick look at the data. You are not expected to understand the code for them, but try to understand what the output means.

### What are the columns?

You'll observe that there are some extra columns now. 

The **id columns** are the columns that uniquely identify an observation (sometimes called a "case" instead of "observation"). ^[The id columns are analogous to *independent variables*, except that that terminology often implies that we're running an experiment, which is not necessarily the case.]
In Homework 1, we only had one id column, `date`, because we had one observation for each date. (Consider why `day_of_week` and `workingday` would not be considered *id* columns.)

The dataset for this homework has three id columns:

* `date`: as before
* `hour`: hour of the day (0 through 23)
* `rider_type`: `registered` or `casual` (see below)

The two additional id columns mean that we've now broken down the data by `hour` of the day and rider type (`rider_type`). 
Some riders have `registered` for a Capital Bikeshare membership to get better rates.
Other riders just bought a single trip or short-term pass, so we call them `casual` riders.

```{marginfigure}
According to the [source data](https://www.capitalbikeshare.com/system-data), "casual" riders include: Single Trip, 24-Hour Pass, 3-Day Pass or 5-Day Pass
```

So: each row is the count of how many *rides* were completed in a given hour by a given type of rider. For each row, we have the following observed variables:^[These are analogous to *dependent variables*, but again that may imply too much.]

* `rides`: the number of rides by that type of rider
* `season`: Winter, Spring, sUmmer, or Fall
* `year`: 0 for 2011, 1 for 2012.
* `holiday`: 0 for ordinary days, 1 for holidays
* `workingday`: either "weekday" or "weekend" (where "weekend" includes holidays too).
* `day_of_week`: an integer between 0 and 6 inclusive. In this exercise, you will decode which number represents Monday, etc.
* `temp`: the average temperature that day, scaled so that the lowest temperature in the dataset is 0 and the highest is 1.
* `feels_like`: the "feels-like" temperature, also scaled to range from 0 to 1.
* `humidity`: relative humidity, also scaled 0 to 1.
* `windspeed`: wind speed, also scaled 0 to 1


## Exercise 1: Label days of week

The data set uses the integers 0 through 6 to label days of the week. It does not
document, however, what 0 means, or what 6 means. If we want to make understandable
plots, we should label these days of the week.

First, *look at the data* to figure out what a `day_of_week` of 0 means. Then what a `day_of_week` of 1 means.
You may find it helpful to refer to a calendar.

Once you've figured it out:

1. Write a brief description of what the mapping is and your evidence that you got it correct.
2. Fill in the blank in the given code block to give labels to the weekdays. Use abbreviations ("Wed", "Fri", ...).

```{r label-weekdays, include=FALSE}
daily_rides <- daily_rides %>%
  mutate(day_of_week = factor(day_of_week, levels = c(0, 1, 2, 3, 4, 5, 6), labels = c("Sun", "Mon", "Tue", "Wed", "Thur", "Fri", "Sat")))
```

## Exercise 2: Describe a row

Write a succinct English sentence that conveys the same information as the first row in the data frame.
*Only consider the `date`, `rider_type`, `rides`, `holiday`, `workingday`, and `day_of_week` columns.*

## Exercise 3: Rides by date, by rider type

Make a scatterplot of the number of rides by date, broken down by type of rider. Tips:

- Refer to your Homework 1 solution for a very similar (but not identical) plot.
- Make the points smaller (`size` of 1) and partially transparent (`alpha` of 0.5) to reduce *overplotting*.
- Your plot should be fully labeled.

Here is one possibility:

```{r}
ggplot(daily_rides, aes(x = date, y = rides, color = rider_type)) +
  geom_point(alpha = 0.5, size = 1) +
  geom_smooth() +
  labs(x = "Date", y = "Number of rides", color = "Type of Rider", title = "Rides by date", subtitle = "for casual and registered riders")
```

Write a brief interpretation of this plot.

## Exercise 4: Mapping vs Faceting

Now let's look at workdays vs weekends (and holidays). This is encoded in the `workingday` variable, like in Homework 1.

First, make a scatterplot like in Exercise 3, but map `shape` to `workingday`.

```{r}
ggplot(daily_rides, aes(x = date, y = rides, color = rider_type, shape = workingday)) +
  geom_point(alpha = 0.5, size = 1) +
  geom_smooth() +
  labs(x = "Date", y = "Number of rides", color = "Type of Rider", title = "Rides by Date", shape = "Type of Day")
```

## Faceting

```{r}
ggplot(hourly_rides, aes(x = date, y = rides, color = rider_type)) +
  geom_point(alpha = 0.2) +
  facet_wrap(vars(workingday), nrow = 2)
```

```{r}
ggplot(hourly_rides, aes(x = date, y = rides, color = workingday)) +
  geom_point(alpha = 0.2) +
  facet_wrap(vars(rider_type), nrow = 2)
```

```{r}
ggplot(hourly_rides, aes(x = date, y = rides, color = workingday)) +
  geom_point(alpha = 0.2) +
  facet_grid(cols = vars(rider_type), rows = vars(hour))
```

## univariate plot

First, make the histogram bars blue.

```{r}
ggplot(hourly_rides, aes(x = rides)) + geom_histogram(binwidth = 50, fill = 'blue')
```

Now, color the bar according to whether it's weekend or workday.

```{r}
ggplot(hourly_rides, aes(x = rides, fill = workingday)) + geom_histogram(binwidth = 50)
```


```{r}
ggplot(hourly_rides, aes(x = date, y = season)) + geom_point()
```


```{r}
ggplot(hourly_rides, aes(x = temp, y = season)) + geom_boxplot()
```




# Exercise 2: Overall counts by day of week

The following code (also in your template) totals the number of rides by each
member by each day of week. We will be learning how to write code like this
next week:

```{r ride-totals}
ride_totals <- hourly_rides %>%
  group_by(rider_type, day_of_week) %>% 
  summarize(rides = mean(rides))
```

Make the following plot. (You may choose to refine it, but a plot like this is acceptable.)

```{r ride-totals-by-weekday, echo=FALSE, fig.fullwidth = TRUE}
ggplot(ride_totals, aes(x = day_of_week, fill = rider_type, y = rides)) +
  geom_col(position = position_dodge()) +
  labs(y = "Rides", x = "Day of Week", fill = "Type of Rider")
```

Hints:

* Use `geom_col`.
* To put the bars side-by-side instead of stacked, use `position = position_dodge()`.


# Exercise 3: Distribution of Rides by Weekday

The bar plot shows a single number for each weekday. Refine the graph to show
the *distribution* of rides each hour.

To do this:

* Use the original `hourly_rides` dataset.
* Use a box plot

Your plot might look something like this:

```{r ride-distributions-by-weekday, echo=FALSE}
hourly_rides %>%
  ggplot(aes(x = day_of_week, y = rides, fill = rider_type)) +
  geom_boxplot() +
  labs(y = "Rides", x = "Day of Week", fill = "Type of Rider")
```

# Exercise 4: Distribution of Rides by Hour

Instead of variation by day of week, plot the variation by hour, distinguishing the type of rider and between working days and weekends. Details:

* The x axis should be hour. (Feel free to try flipping the axes too.)
* Use side-by-side box plots for casual and registered members.
* Label the plot.
* Use `facet_wrap` to make separate graphs for working days and weekends. Try different settings for the scales and number of rows / columns until you find a layout that looks good. (Read the help page for `facet_wrap` to find out what arguments to use.)
* Set the `fig.height` and/or `fig.width` cell options to make a comfortably sized graph.

```{r ride-distributions-by-hour, fig.height=10, fig.width=10, eval=FALSE, include=FALSE}
ggplot(hourly_rides, aes(x = hour, y = rides, fill = rider_type)) + 
  geom_boxplot() +
  facet_wrap(~ workingday, scales = "free_y", nrow = 2)
```

# Exercise 5: Overplotting

The following plot attempts to show how temperature affects ridership. However, it suffers
from *overplotting*.

```{r rides-by-temp-overplot}
ggplot(hourly_rides, aes(x = temp, y = rides)) +
  geom_point() +
  facet_wrap(~ rider_type) + 
  labs(x = "Normalized temperature", y = "Rides")
```

Pick at least two of the following methods to try to fix the overplotting issue. Which do you prefer, and why? Include the code and results of both plots you tried.

* `geom_density_2d`
* Making points transparent (`alpha = 0.05`) or very small (`shape = '.'`)
* binning the x axis (`group = cut_width(temp, ___)`)
* any other approach you want to try!

Note: you don't need to make proper labels for these plots, though it's always nice.

```{r binned-x, include=FALSE}
hourly_rides %>%
  ggplot(aes(x = temp, y = rides, group = cut_width(temp, .1))) +
    geom_boxplot() +
    facet_wrap(~ rider_type)
```
