---
title: "Lab 07: Making Predictive Models"
author: "TEAM"
date: "DATE"
output: github_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
 options(scipen = 5) # encourage metrics to print in fixed-point notation
```

## Load data

```{r load-and-subset-data}
data(ames, package = "modeldata")
ames <- ames %>% 
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal")
```

After filtering, we are left with data for `r nrow(ames)` home sales.

## Exploratory analysis

```{r eda-living-area}
ggplot(ames, aes(x = Gr_Liv_Area, y = Sale_Price)) +
  geom_point(size = .1, alpha = .25) +
  labs(x = "above-grade living area (ft^2)", y = "Sale Price ($)")
```


## Modeling

### Hold out some data to use for validation

```{r train-test-split}
# Seed the random number generator
set.seed(10)
# Split our data randomly
ames_split <- initial_split(ames, prop = 2/3)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

### Specify the model

We want a linear regression, which we'll fit with the `lm` ("linear model") engine that's built-in to R.

```{r model-spec}
my_model_spec <- 
  parsnip::linear_reg() %>% 
  set_mode("regression") %>% 
  set_engine("lm")
```

### Fit the model on data

```{r}
my_trained_model <-
  my_model_spec %>% 
  fit(Sale_Price ~ Lot_Area, data = ames_train)
my_trained_model
```

* What **coefficients** do you see?
* Try re-running the `fit`. Do the coefficients change?
* Try re-running the `initial_split` and then the `fit`. Do the coefficients change?
* Try changing the seed, then re-running. What changes?

### How well did it do?

On data it saw already (the **training set**):

```{r}
train_predictions <- 
  my_trained_model %>% 
    predict(ames_train)
train_predictions
```
* How many predictions did it make? Why might that make sense?


```{r}
train_predictions %>% 
  bind_cols(ames_train) %>% # Put back the original data
  yardstick::metrics(truth = Sale_Price, estimate = .pred)
```

* Why does this table have 3 rows? What does each row mean?
* Ignore the ".estimator" column for now.

### How well did it do... on *unseen* data?


```{r}
my_trained_model %>% 
  predict(ames_test) %>% 
  bind_cols(ames_test) %>% 
  metrics(truth = Sale_Price, estimate = .pred)
```

### Is that good?

Finish the sentence: "On average, this model's predictions were off by about $___".

Can we do better?

* What data about a house did this model use to make its prediction?
* What kinds of computations could it do to that data?

## More features, part 1

Try using `Lot_Area`, 

```{r}
my_trained_model <- my_model_spec %>% 
  fit(Sale_Price ~ Lot_Area + Gr_Liv_Area + Full_Bath, data = ames_train) #<<
```

```{r}
my_trained_model %>% 
  predict(ames_test) %>% 
  bind_cols(ames_test) %>% 
  metrics(truth = Sale_Price, estimate = .pred)
```

### LOTS more features!

Here's how to get a list of *all* the numeric features in formula-like format:

```{r}
ames %>% select(where(is.numeric), -Sale_Price) %>% names() %>% paste0(collapse = " + ")
```


```{r}
my_trained_model <- my_model_spec %>% 
  fit(Sale_Price ~ Lot_Frontage + Lot_Area + Year_Built + Year_Remod_Add + Mas_Vnr_Area + BsmtFin_SF_1 + BsmtFin_SF_2 + Bsmt_Unf_SF + Total_Bsmt_SF + First_Flr_SF + Second_Flr_SF + Gr_Liv_Area + Bsmt_Full_Bath + Bsmt_Half_Bath + Full_Bath + Half_Bath + Bedroom_AbvGr + Kitchen_AbvGr + TotRms_AbvGrd + Fireplaces + Garage_Cars + Garage_Area + Wood_Deck_SF + Open_Porch_SF + Enclosed_Porch + Three_season_porch + Screen_Porch + Pool_Area + Misc_Val + Mo_Sold + Year_Sold + Longitude + Latitude, data = ames_train)
```

```{r}
my_trained_model %>% 
  predict(ames_test) %>% 
  bind_cols(ames_test) %>% 
  metrics(truth = Sale_Price, estimate = .pred)
```





