---
title: "Lab 07: Making Predictive Models"
output: 
  tufte::tufte_html:
    css: ../lab.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(glue)
 options(scipen = 5) # encourage metrics to print in fixed-point notation
```

## Goals

* Apply a basic `tidymodels` workflow to fit and evaluate a linear regression model
* Practice *model validation* on held-out data
* Practice more visualization (as always!)

## Getting started

Find your repo in [our GitHub organization](https://github.com/Calvin-DS202-FA20/)
and clone it as usual.

### Optional, experimental workflow

The RStudio server lets multiple people edit a single project at the same time!
This can be confusing since it's different from collaborating using GitHub, so
try it only if all of your team is game to try:

1. The person whose name appears first on the Attendance sheet (or someone else if needed)
  clones the team's repo on the RStudio Server.
2. They 

We'll be using `tidymodels`:

```{r message=FALSE}
library(tidymodels)
```

## Load data

We subset the data as [De Cock](http://jse.amstat.org/v19n3/decock.pdf) suggests.
Again, see [Data dictionary](http://jse.amstat.org/v19n3/decock/DataDocumentation.txt)

```{r}
data(ames, package = "modeldata")
ames <- ames %>% 
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal")
nrow(ames)
```

## Exploratory analysis

```{r eda-living-area}
ggplot(ames, aes(x = Gr_Liv_Area, y = Sale_Price, color = Fireplaces > 0)) +
  geom_point()
```

```{r eda-several-predictors}
ames %>% select(Sale_Price, Gr_Liv_Area, Lot_Area, Full_Bath, Half_Bath, Fireplaces) %>% 
  pivot_longer(-c(Sale_Price, Fireplaces), names_to = "predictor", values_to = "predictor_value") %>% 
  ggplot(aes(x = predictor_value, y = Sale_Price, color = Fireplaces > 0)) + geom_point() +
  facet_wrap(vars(predictor), scales = "free") + theme_bw()
```

## Modeling

### Hold out some data to use for validation

```{r}
set.seed(10)
ames_split <- initial_split(ames, prop = 1/2)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
glue("Using {nrow(ames_train)} sales to train, {nrow(ames_test)} to test")
```

* What does `set.seed` do? Why might it be a good idea?
* Why might 3/4 be a reasonable proportion?

### Specify the model

We want a linear regression, which we'll fit with the `lm` ("linear model") engine that's built-in to R.

```{r}
# my_model_spec <-
#   parsnip::rand_forest() %>%
#   set_mode("regression") %>% 
#   set_engine("ranger")
my_model_spec <- 
  parsnip::linear_reg() %>% 
  set_mode("regression") %>% 
  set_engine("lm")
```

### Fit the model on data

```{r}
my_trained_model <-
  my_model_spec %>% 
  fit(Sale_Price ~ Lot_Area, data = ames_train)
my_trained_model
```

* What **coefficients** do you see?
* Try re-running the `fit`. Do the coefficients change?
* Try re-running the `initial_split` and then the `fit`. Do the coefficients change?
* Try changing the seed, then re-running. What changes?

### How well did it do?

On data it saw already (the **training set**):

```{r}
train_predictions <- 
  my_trained_model %>% 
    predict(ames_train)
train_predictions
```
* How many predictions did it make? Why might that make sense?


```{r}
train_predictions %>% 
  bind_cols(ames_train) %>% # Put back the original data
  yardstick::metrics(truth = Sale_Price, estimate = .pred)
```

* Why does this table have 3 rows? What does each row mean?
* Ignore the ".estimator" column for now.

### How well did it do... on *unseen* data?


```{r}
my_trained_model %>% 
  predict(ames_test) %>% 
  bind_cols(ames_test) %>% 
  metrics(truth = Sale_Price, estimate = .pred)
```

### Is that good?

Finish the sentence: "On average, this model's predictions were off by about $___".

Can we do better?

* What data about a house did this model use to make its prediction?
* What kinds of computations could it do to that data?

## More features, part 1

Try using `Lot_Area`, 

```{r}
my_trained_model <- my_model_spec %>% 
  fit(Sale_Price ~ Lot_Area + Gr_Liv_Area + Full_Bath, data = ames_train) #<<
```

```{r}
my_trained_model %>% 
  predict(ames_test) %>% 
  bind_cols(ames_test) %>% 
  metrics(truth = Sale_Price, estimate = .pred)
```

### LOTS more features!

Here's how to get a list of *all* the numeric features in formula-like format:

```{r}
ames %>% select(where(is.numeric), -Sale_Price) %>% names() %>% paste0(collapse = " + ")
```


```{r}
my_trained_model <- my_model_spec %>% 
  fit(Sale_Price ~ Lot_Frontage + Lot_Area + Year_Built + Year_Remod_Add + Mas_Vnr_Area + BsmtFin_SF_1 + BsmtFin_SF_2 + Bsmt_Unf_SF + Total_Bsmt_SF + First_Flr_SF + Second_Flr_SF + Gr_Liv_Area + Bsmt_Full_Bath + Bsmt_Half_Bath + Full_Bath + Half_Bath + Bedroom_AbvGr + Kitchen_AbvGr + TotRms_AbvGrd + Fireplaces + Garage_Cars + Garage_Area + Wood_Deck_SF + Open_Porch_SF + Enclosed_Porch + Three_season_porch + Screen_Porch + Pool_Area + Misc_Val + Mo_Sold + Year_Sold + Longitude + Latitude, data = ames_train)
```

```{r}
my_trained_model %>% 
  predict(ames_test) %>% 
  bind_cols(ames_test) %>% 
  metrics(truth = Sale_Price, estimate = .pred)
```





