---
title: "Midterm 1"
author: "K Arnold"
date: "10/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(nycflights13)
```

## Objectives

- Relate the structure of a dataset to the structure of a visualization of it
- Analyze a visualization in terms of the grammar of graphics.
- Predict the effects of fundamental data manipulation operations
- Compose a sequence of data manipulation steps to achieve a desired data manipulation operation


## Exercises

Consider the following two plots.

TODO describe this data, label the plots

```{r}
mean_delays_by_hr <- nycflights13::flights %>% 
  filter(!is.na(dep_delay)) %>% 
  group_by(origin, hour) %>% 
  summarize(mean_delay = mean(dep_delay, na.rm = TRUE), .groups = "drop")

mean_delays_by_hr %>% 
  ggplot(aes(x = hour, y = mean_delay, color = origin)) +
    geom_point()

mean_delays_by_hr %>% 
  ggplot(aes(x = origin, y = mean_delay, fill = factor(hour))) +
    geom_col(position = "dodge")
```

1. What type of plot is plot A?

2. What type of plot is plot B?

3. Make a table of the aesthetic mappings for plot A.

4. Make a table of the aesthetic mappings for plot B.

5. Suppose plot A was constructed by supplying `tableA` to a `ggplot` command, and that nothing fancy happens in the `ggplot` (no `geom`s that count rows, no `stat_*` commands, etc.).

  a. Give a plausible example of the first 3 rows of `tableA` (i.e., `head(tableA, 3)`).
  
  b. How many rows do you expect `tableA` to have? Write (but do not evaluate) the multiplication expression that you would enter into a simple calculator to give the this result.

6. Repeat problem 5 for plot B.

7. Give an example of something that you can notice about the data (i.e., a comparison you can make) *more easily in plot A than in plot B*.

  a. What is the specific observation you make? (e.g., "The mean delay for XXX is higher for YYY than ZZZ")
  
  b. What about the design of the visualization makes that comparison easier in plot A?

8. Repeat problem 7, but for plot B instead.

9. Identify a change to make to plot A that would improve it.

  a. Circle or draw an arrow to the part of the plot that you would change.
  
  b. Sketch what that part of the plot would look like after your change.
  
10. Repeat problem 9 for plot B.

### Contrast these two plots?

One with inappropriate aesthetic mapping, e.g., color for date.

### Design a vis

- Sketch a *univariate* visualization of one variable of this dataset. Include all needed labels and guides.
  - Name of the plot type.
  - Describe what variable is mapped to what aesthetic.
- Sketch a *bivariate* (or higher dimension) visualization.
  - Name plot type
  - Describe what variable is mapped to what aesthetic.

### Describe structure along a pipeline

- Given a basic group-summarize-arrange-join pipeline, draw the shape of the data, including all column names and a mutually consistent row count.

  
  
