---
title: "Guided Exercise 13 - Text"
output: 
  tufte::tufte_html:
    css: ../ex.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(tidyverse)
theme_set(theme_bw())
options(scipen = 5) # encourage metrics to print in fixed-point notation
options(dplyr.summarise.inform = FALSE) # silence a warning message
```

Humans, created as representatives of a speaking God, do a lot of speaking. So, naturally, text data is everywhere: social media posts, product descriptions, customer surveys, newspaper articles, emails, press releases, etc.

Words is amazingly flexible, but sometimes we want to quantify, which entails reducing that flexibility to overly simplistic numbers.
Keeping in mind that trade-off, this exercise is about equipping you with a few tools to turn text into tabular data.
The tools we'll see today are:

- *regular expressions* (**regex**): a little programming language used for looking for specific kinds of things in text strings. For example, the regular expression `th[eo]se` matches either "these" or "those" (and nothing else)
- *web scraping*: a way of pulling text (or other data) out of loosely-structured web pages

Learning objectives:

- Predict how a regular expression will match a given text
- Write regular expressions to perform extraction and separation tasks
- Scrape data from web pages

## Getting started

Pull your portfolio repo as usual.

We'll be working through exactly the examples in [Chapter 19](https://mdsr-book.github.io/mdsr2e/ch-text.html) of our textbook, so you might want to have that handy.
You may also want to refer to the [Strings chapter in *R for Data Science*](https://r4ds.had.co.nz/strings.html) for more details.

## `separate`

First, let's hit one of the most common things you need to do with string data: separate a string into several pieces. So: **do the `separate()` section of the [Tidy Your Data primer](https://rstudio.cloud/learn/primers/4.2)**. (Don't worry about `unite()`, and come back later to the Case Study if you want more practice with a cool wrangling puzzle.)

> Note: the tutorial uses the old name of `pivot_longer`, which was `gather`.


## Macbeth

```{r}
macbeth_url <- "http://www.gutenberg.org/cache/epub/1129/pg1129.txt"
#macbeth_raw <- read_file(macbeth_url)
data(Macbeth_raw, package = "mdsr")
```

We'll start with a data frame that has a row for each line in Macbeth.

```{r}
macbeth <- Macbeth_raw %>% 
  stringi::stri_split_lines() %>%
  pluck(1) %>% 
  enframe("line_num", "line")
macbeth
```

The `str_detect` function tells whether or not a line contains a pattern.

```{r}
pattern <- "MACBETH"
macbeth %>% 
  mutate(matches = str_detect(line, pattern)) %>% 
  filter(matches)
```

1. How many lines contain "MACBETH"? How many lines contain "MACDUFF"?

The `.` character matches any single character.

```{r}
pattern <- "MAC."
macbeth %>% 
  mutate(matches = str_detect(line, pattern)) %>% 
  filter(matches)
```

The `extract` function (from `tidyr`) pulls out what matched:

```{r}
macbeth %>% 
  extract(line, into = "what_matched", regex = )
```

