---
title: "Making Inferences about Features"
author: "K Arnold"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: FALSE
    lib_dir: libs
    df_print: paged
    nature:
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
source("../slides-common.R")
slideSetup()
knitr::opts_chunk$set(echo = TRUE)
library(tidymodels)
library(vip)
```

## What's your goal?

* **Predict** unseen labels
  * How much will this house sell for?
  * Does this child have autism?
  * Is this a positive or negative movie review?
* **Infer** relationships between features and labels
  * How much does home size affect price?
  * Is DNA methylation a marker of autism?
  * Does "sick" indicate a positive or negative review?
* Understand the **causal** effect of interventions
  * How much will building an addition increase the price of my home?
  * Will antioxidants prevent autism?
  * Will cutting this scene make my movie get better reviews?

---

## Techniques for Inference

* Classical statistical tests (2-sample t tests, chi squared tests, ANOVA, ...)
* Variable importance plots
* Benefit of adding each feature

---

## First, a caution

(Palmer Penguins example of Simpson's Paradox)

---

## Variable Importance Plots

```{r load-data, echo=FALSE}
#data(ames, package = "modeldata")
ames <- AmesHousing::make_ames()
ames_all <- ames %>%
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal") %>%
  mutate(across(where(is.integer), as.double)) %>%
  mutate(Sale_Price = Sale_Price / 1000)
rm(ames)
```

```{r train-test-split, echo=FALSE}
set.seed(10) # Seed the random number generator
ames_split <- initial_split(ames_all, prop = 2 / 3)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

```{r}
regresion_workflow <- workflow() %>%  add_model(decision_tree(mode = "regression") %>% set_engine('rpart')) 
```


```{r}
model <- regresion_workflow %>% 
  add_recipe(recipe(Sale_Price ~ ., data = ames_train)) %>% 
  fit(data = ames_train)
```

```{r tree-vip}
model %>% 
  pull_workflow_fit() %>% 
  vip::vip(num_features = 20L)
```

---

## How much does it help to have a feature in?

```{r make-resamples}
set.seed(20201118)
resamples <- vfold_cv(ames_train, v = 10)
```

```{r tree-all-features}
regresion_workflow %>% 
  add_recipe(recipe(Sale_Price ~ ., data = ames_train)) %>% 
  fit_resamples(resamples = resamples, metrics = metric_set(mae, rmse)) %>% collect_metrics()
```



```{r tree-without-qual}
regresion_workflow %>% 
  add_recipe(recipe(Sale_Price ~ ., data = ames_train) %>% step_rm(ends_with("Qual"))) %>% 
  fit_resamples(resamples = resamples, metrics = metric_set(mae, rmse)) %>% collect_metrics()
```


---

## Appendix: code

.small-code[
```{r get-labels, echo=FALSE}
labels_of_non_echoed_code <- setdiff(
  knitr::all_labels(echo == FALSE),
  c("setup", "get-labels")
)
```

```{r all-code, ref.label=labels_of_non_echoed_code, eval=FALSE}
```
]
