---
title: "HW 3: Bikeshare Aggregation"
output: 
  tufte::tufte_html:
    css: ../hw.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Disable a warning message.
options(dplyr.summarise.inform = FALSE)
```

In this exercise, we'll continue our study of the bike sharing data, but we’re
starting to take off the training wheels: rather than using
pre-wrangled data, you’ll be doing some of the data wrangling yourself!

In this exercise you'll:

* *refine* your skills in creating reproducible documents (R Markdown), graphics (ggplot2), and filtering functions
* *learn* grouping and aggregation functions, and one pivoting operation (`pivot_wider`)
* *see* some ways to work with dates

## Getting started

Go to [our GitHub org](https://github.com/Calvin-DS202-FA20/) and find your `hw03` repo.

(The "GitHub Classroom" approach has stopped working for us.)

As usual, clone the repo and try to knit the template `Rmd`.

The template uses the following packages.

```{r packages, message=FALSE}
library(tidyverse)
library(lubridate)
```


# Load Data

We downloaded the data from the [Capital BikeShare program](https://www.capitalbikeshare.com/system-data). Note the license agreement there. We removed some columns of the data that we aren't using to reduce the size of the dataset you have to download.

The data files are broken up by year. Here we read in the data for both 2011 and 2012, the same two years that you worked with in Homework 1 and 2. The template already includes these lines, so you can just Run All to get the data loaded.

```{r load-data}
# Read the file, specifying data types for each column.
rides_2011 <- read_csv("data/rides_2011.csv.gz", col_types = cols(
  start_time = col_datetime(),
  duration = col_double(),
  member_type = col_factor()
))
rides_2012 <- read_csv("data/rides_2012.csv.gz", col_types = cols(
  start_time = col_datetime(),
  duration = col_double(),
  member_type = col_factor()
))
```


Each row is a *ride*. For each ride we're given the start time of the trip, its duration, and "member type".
According to the website:

> Member Type – Indicates whether user was a "registered" member (Annual Member, 30-Day Member or Day Key Member) or a "casual" rider (Single Trip, 24-Hour Pass, 3-Day Pass or 5-Day Pass)


1. Use the code below to combine the data from these two years into a single data frame called `rides`.
Report some basic information about its shape.

```{r bind-rows}
rides <- bind_rows(
  rides_2011,
  rides_2012
)
```

2. What units is `duration` probably measured in? Make a visualization to justify your answer.


3. Make a table that shows how many rides were taken by members and how many were taken by non-members.

```{marginfigure}
Note that `summarize` and `summarise` are synonyms.
```

Make the same result in two ways: once with `count`, and once with `group_by` and `summarize`.
Your result should look like this (use `knitr::kable()` at the end of your pipe if you want the nice formatting).

```{r count-member-type, echo=FALSE}
rides %>%
  group_by(member_type) %>% 
  summarize(n = n()) %>% 
#  count(member_type) %>%
  knitr::kable()
```

4. Create a dataframe like the one used in Homework 2 that includes the number of rides in each hour. How many rows does this data frame have? How many *should* it have? (Don't try to fix this right now.)

```{marginfigure}
See the `lubridate` [cheat sheet](https://rawgit.com/rstudio/cheatsheets/master/lubridate.pdf) and the ["dates and times" chapter](https://r4ds.had.co.nz/dates-and-times.html) of "R for Data Science".
```

* The `floor_date` function can be used to round dates to the nearest hour (see its `unit` parameter).
* You may use `count`, but also try using `summarize(n = n())`

```{r aggregate-hourly, echo=FALSE}
hourly_rides <- rides %>%
  group_by(start_hour = floor_date(start_time, unit = "hours")) %>% 
  count()
```


5. Break down the result by casual vs registered member. The result should look like:

```{r count-by-hour-and-member-type, echo=FALSE}
hourly_rides <- rides %>%
  group_by(hour = floor_date(start_time, unit = "hours"), member_type) %>% 
  summarize(rides = n())
hourly_rides
```

6. Repeat, aggregating by date instead. 

7. Recreate the "Distribution of Rides by Weekday" plot from Homework 2 using this dataset.

Notes:

* You will need the `wday` function (from the `lubridate` package) to get the day of week. Give it `label = TRUE` to get "Sun", "Mon" instead of numbers.
* Omit the "Unknown" member type.

```{r rides-by-weekday, echo=FALSE}
hourly_rides %>%
  filter(member_type != "Unknown") %>% 
  ggplot(aes(x = wday(hour, label = TRUE), y = rides, fill = member_type)) +
  geom_boxplot() +
  labs(y = "Rides", x = "Day of Week", fill = "Type of Rider")
```

8. Make a tabular summary of this same dataset: what is the median number of rides taken by each type of rider for each day of the week?

* Once you get the basic table, try adding as the last step of your pipeline `pivot_wider(names_from = ___, values_from = ___)`, filling in the blanks appropriately.
* End your pipeline with `knitr::kable()` for a nice-looking table.

The result should look like:

```{r rides-by-weekday-table, echo=FALSE}
hourly_rides %>% 
  filter(member_type != "Unknown") %>% 
  group_by(weekday = wday(hour, label = TRUE), member_type) %>% 
  summarize(median_rides = median(rides)) %>% 
  pivot_wider(names_from = member_type, values_from = median_rides) %>% 
  knitr::kable()
```


9. Recreate the "Distribution of Rides by Hour" plot using this dataset.

* Note: the hw2 dataset had marked some holidays as non-working days. For this plot, let's ignore
  that and just assume that Saturday and Sunday are the only non-working days.
* You will need these functions:
  * `lubridate::hour`
  * `as_factor`
  * `if_else` or `case_when`.

```{r ride-distributions-by-hour, fig.height=10, fig.width=10, echo=FALSE}
hourly_rides %>% 
  filter(member_type != "Unknown") %>% 
  mutate(is_weekend = if_else(
    wday(hour, label = TRUE) %in% c("Sat", "Sun"),
    "Weekend", "Weekday")) %>% 
  ggplot(aes(x = as_factor(lubridate::hour(hour)), y = rides, fill = member_type)) + 
    geom_boxplot() +
    facet_wrap(~ is_weekend, scales = "free_y", nrow = 2)
```



## Appendix: Preprocessing

This is the code that I used to download and preprocess the dataset.

```{r eval=FALSE, include=FALSE}
datafiles <- tribble(
  ~ url, ~destFile,
  "https://s3.amazonaws.com/capitalbikeshare-data/2011-capitalbikeshare-tripdata.zip", "rides_2011.csv",
  "https://s3.amazonaws.com/capitalbikeshare-data/2012-capitalbikeshare-tripdata.zip", "rides_2012.csv",
)
```


```{r eval=FALSE}
preprocFile <- function(url, destFile) {
  url %>%
    pin() %>%
    map_dfr( ~ read_csv(
      .,
      col_types = cols_only(
        Duration = col_double(),
        `Start date` = col_datetime(format = ""),
        `Member type` = col_character()
      )
    )) %>%
    select(start_time = `Start date`,
         duration = `Duration`,
         member_type = `Member type`) %>% 
    write_csv(destFile)
  system2("gzip", c("-9v", destFile))
}
preprocFile(
  "https://s3.amazonaws.com/capitalbikeshare-data/2011-capitalbikeshare-tripdata.zip",
  "rides_2011.csv"
)
preprocFile(
  "https://s3.amazonaws.com/capitalbikeshare-data/2012-capitalbikeshare-tripdata.zip",
  "rides_2012.csv"
)
```

