---
title: "Cooking with Models"
author: "K Arnold"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: FALSE
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
source("../slides-common.R")
slideSetup()
knitr::opts_chunk$set(echo = TRUE)
library(glue)
```

## HW 06 notes

* Get as *close to the source* as you can
  * e.g., "Our World in Data" is great, but often data is pre-wrangled. See "Source" tab.
* *tidy data tip*: Try expressing your data as a bunch of simple sentences, one per observation.
  * On January 22, the number of Covid cases in Afghanistan was 0.
  * On January 23, the number of Covid cases in Afghanistan was 0.
  * ...
* For bar charts, use `geom_col` (not `geom_bar`) unless you know exactly what you're doing.

---

## Objectives

* What are the basic steps in training and validating any predictive model?
* Why is each step important?
* How can we use the `tidymodels` ecosystem to train and validate a linear model?

---

## A Grammar of Predictive Modeling

1. What problem are you trying to solve? (predict *what*, based on *what*?)
2. Which type of model do you want to use?
3. How do you want to transform your data before giving it the model?
4. How will you measure how well your models work?
5. How will you split up your data between training and validation?

```{r}
library(tidymodels)
```

Basic steps:

* `parsnip`: **Specify** and **train** the model you want
* `recipes`: **Prepare** the data
* `rsample`: **Sample** training and validation data
* `yardstick`: Compute **metrics** for performance

---

## Example data

We subset the data as De Cock suggests:

```{r}
ames_raw <- AmesHousing::make_ames()
ames <- ames_raw %>% 
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal")
nrow(ames)
names(ames)
#ggplot(ames, aes(x = Gr_Liv_Area, y = Sale_Price, color = Fireplaces > 0))+geom_point()
```

---

## Example, without validation

Specify the model:

```{r}
my_model_spec <- 
  parsnip::linear_reg() %>% 
  set_engine("lm")
```

Train it ("fit") on data:

```{r}
my_trained_model <- 
  my_model_spec %>% 
  fit(Sale_Price ~ Lot_Area, data = ames)
```

Predict on new data:

```{r include=FALSE}
augment.model_fit <- function(x, ...) {
  if ("new_data" %in% names(list(...))) {stop("Use newdata, not new_data")}
  augment(x$fit, ...)
}
```


```{r}
tibble(Lot_Area = c(32000, 12000), Sale_Price = c(239000, 185000)) %>%
  predict(my_trained_model, new_data = .)  
  #augment(my_trained_model, newdata = .)
```

---

## Example, with validation

Hold out some data to use for validation:

```{r}
set.seed(10)
ames_split <- initial_split(ames, prop = 3/4)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
glue("Using {nrow(ames_train)} sales to train, {nrow(ames_test)} to test")
```


Specify and train the model **on the training set**:

```{r}
my_trained_model <- 
  my_model_spec %>% 
  fit(Sale_Price ~ Lot_Area + Gr_Liv_Area + Full_Bath, data = ames_train) #<<
```

Make predictions:

```{r}
train_predictions <- 
  my_trained_model %>% 
    predict(ames_train) %>% 
    bind_cols(ames_train) # Put back the original data
```

```{r}
train_predictions %>% 
  yardstick::metrics(truth = Sale_Price, estimate = .pred)
```

```{r}
my_trained_model %>% 
  predict(ames_test) %>% 
  bind_cols(ames_test) %>% 
  metrics(truth = Sale_Price, estimate = .pred)
```


---

## Types of models

* Linear models
  * ordinary least-squares (OLS)
  * Lasso, Ridge, etc.: penalize large coefficients
  * Generalized Linear Models: outputs get transformed
    * Logistic Regression (also Support Vector Machine): transform output to *score* for each class
* Decision Lists and Trees
  * extension: Random Forests
* Neural Networks: layered combinations of the above
* many, many more

---

## Which variables mean what?

The *formula interface*

don't get confused: they "forgot" the coefficients!

But this works for specifying even models that aren't linear.

This is `+` like `ggplot` or like "determination + perseverance + grit = success",
not like "2x + y = 4".

---


