---
title: "Cooking with Models"
author: "K Arnold"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: FALSE
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
source("../slides-common.R")
slideSetup()
knitr::opts_chunk$set(echo = TRUE)
library(glue)
```

## HW 06 notes

* Get as *close to the source* as you can
  * e.g., "Our World in Data" is great, but often data is pre-wrangled. See "Source" tab.
* *tidy data tip*: Try expressing your data as a bunch of simple sentences, one per observation.
  * On January 22, the number of Covid cases in Afghanistan was 0.
  * On January 23, the number of Covid cases in Afghanistan was 0.
  * ...
* For bar charts, use `geom_col` (not `geom_bar`) unless you know exactly what you're doing.
* If you're having trouble getting data, talk with me!

---

## Q&A

> Should I look at (root) mean squared error, mean absolute error, or what?

Analogous question: Would I use the mean or the median to summarize the errors?

* If the model is mostly good but makes a few large errors, is that bad (use MSE!), or
  does it mean we should probably ignore those points as outliers (use MAE!).
* **Median** minimizes MAE. **Mean** minimizes MSE.
* Most *models* try to minimize MSE. (but that doesn't mean *you* have to.)

.small[A nice discussion, including concrete examples, [here](https://medium.com/human-in-a-machine-world/mae-and-rmse-which-metric-is-better-e60ac3bde13d).]

---

## Q&A

> We found coefficients by guess-and-check... is there a better way?

**Practically**: machine learning / stats software include efficient algorithms
for exactly this. But it can still take weeks on supercomputers for big
models and data.

**Mathematically**: Most common algorithm is *gradient descent*:

1. From your data, randomly pick a batch of a few observations.
2. Put those X's through your model, tracing the computations on the way.
3. Compute error (e.g., MSE) on that batch.
4. Compute *what small change to each coefficient would have reduced error*?
5. Make all of those small changes, repeat.


---

## Objectives

* What are the basic steps in training and validating any predictive model?
* Why is each step important?
* How can we use the `tidymodels` ecosystem to train and validate a linear model?

---

```{r echo=FALSE}
include_graphics("img/modeling-process-FES.svg")
```

.small[Source: [Feature Engineering and Selection ch1](https://bookdown.org/max/FES/important-concepts.html)]

---

## Predictive Modeling Workflow

Preliminaries:

1. **Define the problem**: predict *what*, based on *what*? What *metrics* will indicate success? (Measure success in multiple ways!)
2. **Explore your data** (EDA): understand its structure, make lots of plots

Modeling:

1. **Pick a model**: Which type(s) of models are appropriate for task and data?
2. **Transform the data** as needed by the model ("feature engineering", preprocessing", "recipe")
3. **Split the data** to allow for validation.
4. **Fit and evaluate the model**
5. **Tune**: adjust model hyperparameters
6. **Analyze model errors** and refine all earlier steps

---


```{r}
library(tidymodels)
```

Packages:

* `parsnip`: **Specify** and **train** the model you want
* `recipes`: **Prepare** the data
* `rsample`: **Split** data into training and validation
* `yardstick`: Compute **metrics** for performance
* `tune`: Helps you set the dials.

---

## Where to find documentation

* TidyModels website: [Getting Started](https://www.tidymodels.org/start/),
  [vignettes](https://www.tidymodels.org/learn/)
* [Tidy Modeling with R](https://www.tmwr.org/) book (work in progress)

.small[
Some others:

* <https://rviews.rstudio.com/2019/06/19/a-gentle-intro-to-tidymodels/>
* <https://juliasilge.com/blog/intro-tidymodels/>
]

---

## Example data

We subset the data as [De Cock](http://jse.amstat.org/v19n3/decock.pdf) suggests.
Again, see [Data dictionary](http://jse.amstat.org/v19n3/decock/DataDocumentation.txt)

```{r}
ames_raw <- AmesHousing::make_ames()
ames <- ames_raw %>% 
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal")
nrow(ames)
names(ames)
#ggplot(ames, aes(x = Gr_Liv_Area, y = Sale_Price, color = Fireplaces > 0))+geom_point()
```

---

## Example, without validation

Specify the model:

```{r}
my_model_spec <-  parsnip::linear_reg() %>% #<<
  set_engine("lm")
```

Train it ("fit") on data:

```{r}
my_trained_model <- my_model_spec %>% 
  fit(Sale_Price ~ Lot_Area, data = ames)
```

Predict on new data:

```{r include=FALSE}
augment.model_fit <- function(x, ...) {
  if ("new_data" %in% names(list(...))) {stop("Use newdata, not new_data")}
  augment(x$fit, ...)
}
```


```{r}
tibble(Lot_Area = c(32000, 12000), Sale_Price = c(239000, 185000)) %>%
  predict(my_trained_model, new_data = .)  
```

---

## Example, with validation

Hold out some data to use for validation:

```{r}
set.seed(10)
ames_split <- initial_split(ames, prop = 3/4)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
glue("Using {nrow(ames_train)} sales to train, {nrow(ames_test)} to test")
```


Specify and train the model **on the training set**:

```{r}
my_trained_model <- 
  my_model_spec %>% 
  fit(Sale_Price ~ Lot_Area + Gr_Liv_Area + Full_Bath, data = ames_train) #<<
```

Make predictions:

```{r}
train_predictions <- 
  my_trained_model %>% 
    predict(ames_train) %>% 
    bind_cols(ames_train) # Put back the original data
```

```{r}
train_predictions %>% 
  yardstick::metrics(truth = Sale_Price, estimate = .pred)
```

```{r}
my_trained_model %>% 
  predict(ames_test) %>% 
  bind_cols(ames_test) %>% 
  metrics(truth = Sale_Price, estimate = .pred)
```


---

## Types of models

* Linear models
  * ordinary least-squares (OLS)
  * Lasso, Ridge, etc.: penalize large coefficients
  * Generalized Linear Models: outputs get transformed
    * Logistic Regression (also Support Vector Machine): transform output to *score* for each class
* Decision Lists and Trees
  * extension: Random Forests
* Neural Networks: layered combinations of the above
* many, many more

---

## Which variables mean what?

The *formula interface*

don't get confused: they "forgot" the coefficients!

But this works for specifying even models that aren't linear.

This is `+` like `ggplot` or like "determination + perseverance + grit = success",
not like "2x + y = 4".

---


