---
title: "Feature Engineering"
author: "K Arnold"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: FALSE
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
source("../slides-common.R")
slideSetup()
knitr::opts_chunk$set(echo = TRUE)
```

## Feedback

Good things:

* Coding together
* Motivating: lab, hw, visualizations, applications

To improve:

* Structure in class
  * Weekly plan is on Moodle
* Lots of different things to do. So simplifying:
  * Check-in quizzes open at beginning of class. Will leave 1 min at end to fill out.
  * Flexible teamwork: each cohort will get 5 repos. Split up as you desire.
  * Goal: no required teamwork outside of class time
* Support outside of class
  * Many students have reached out on Teams at all hours.
  * Feedback / grading

---

## General Hints

* Visualization:
  * What *glyph* represents each *observation*? 
  * What *attributes* ("aesthetics") does that glyph have? (x, y, width, color, ...)
  * What *controls* each aesthetic? ("different political parties have different *y* positions", ...)
* Wrangling
  * What does the *input* look like? (Translate the first row into a data sentence in English.)
  * What does the *output* need to look like? (Again, write a sentence.)
  * What *sequence of steps* needs to happen?
    * filter-groupby-summarize-arrange
    * mutate-groupby-summarize-filter
    * etc.
* Modeling
  * What *quantity* are you trying to predict?
  * What *error measure* will tell you the prediction is good / bad?
  * What *features* can help you make that prediction?

---

## Q&A

> Why don't we just use `lm(y ~ x)` like other stats classes?

* We'll be using many other kinds of models; we're starting with linear models because many people have seen them before.
* `tidymodels` gives a unified interface to lots of different models
* Formulas support some kinds of feature engineering (e.g., `y ~ x1 * x2`) but limited.

---

## Recipes

### Why

Recipes can help us:

* Add expressive power (like conditional logic) to simple models
* Make the model more (or less!) understandable

### What

A **recipe** is a data processing *pipeline* (like `%>%`) with "smart" steps.

---

```{r}
library(tidymodels)
data(ames, package = "modeldata")
ames <- ames %>% 
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal") %>% 
  mutate(across(where(is.integer), as.double))
```

```{r train-test-split}
# Seed the random number generator
set.seed(10)
# Split our data randomly
ames_split <- initial_split(ames, prop = 2/3)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

We'll use one example home from the test set.

```{r example-home}
example_home <- ames_test %>% slice(1)
example_home %>% select(Gr_Liv_Area, Sale_Price)
```


---

```{r}
ames_recipe_1 <- 
  recipe(Sale_Price ~ Gr_Liv_Area + Latitude + Longitude, data = ames_train) %>% 
#  step_range(Gr_Liv_Area, min = 0, max = 1) %>% 
  prep()
ames_recipe_1 %>% summary()
ames_recipe_1 %>% bake(new_data = ames_train)

```

---

## Workflows

`workflow` = `recipe` + `model`

```{r}
workflow1 <- workflow() %>%
  add_model(linear_reg()) %>% 
  add_recipe(ames_recipe_1)
```

Workflows can `fit` and `predict`.

```{r message=FALSE}
fitted_workflow1 <- fit(workflow1, data = ames_train)

fitted_workflow1 %>% predict(example_home)
```

---

```{r unscaled-latlong}
fitted_workflow1 %>% 
  tidy() %>%
  filter(term != "(Intercept)") %>% 
  ggplot(aes(x = estimate, y = fct_reorder(term, estimate))) + geom_col()
```


---

```{r}
ames_recipe_2 <- 
  recipe(Sale_Price ~ Gr_Liv_Area + Latitude + Longitude, data = ames_train) %>% 
  step_range(all_numeric(), min = 0, max = 1) %>% #<<
  prep()
ames_recipe_2 %>% bake(new_data = ames_train)

workflow2 <- workflow() %>%
  add_model(linear_reg()) %>% 
  add_recipe(ames_recipe_2)

fitted_workflow2 <- fit(workflow2, data = ames_train)
```

```{r scaled-latlong}
fitted_workflow2 %>% 
  tidy() %>%
  filter(term != "(Intercept)") %>% 
  ggplot(aes(x = estimate, y = fct_reorder(term, estimate))) + geom_col()
```



---

## What can a linear model do?

* *Add* up terms.
* Each term: *multiply* a number by a constant.

.pull-left[
```{r coefs-from-fitted-model}
intercept <- 15793
coef_living_area <- 108
```


```{r}
intercept + coef_living_area * 1610
```
]

.pull-right[
```{r price-vs-liv-area, out.width="100%"}
ggplot(ames, aes(x = Gr_Liv_Area, y = Sale_Price)) +
  geom_point() +
  geom_hline(yintercept = example_home$Sale_Price, color = "red") +
  geom_vline(xintercept = example_home$Gr_Liv_Area, color = "red") +
  geom_point(data = example_home, color = 'red', size = 5) +
  geom_function(fun = function(x) intercept + coef_living_area * x, color = "blue")
```
]
---

## Conditional Logic: Simple Conditions

How could a linear model 

