---
title: "Python Data Wrangling and Classification"
author: "K Arnold"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: FALSE
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
source("../slides-common.R")
library(readr)
slideSetup(mark_languages = TRUE)
knitr::opts_chunk$set(echo = TRUE)
theme_set(theme_bw())
```

## Example Dataset: Titanic Passengers

* <https://www.openml.org/d/40945>
* <http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/titanic.html>
* <http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/titanic3info.txt>
* <https://www.encyclopedia-titanica.org/>

Download the data, if we don't have it alreday:

```{r download-data}
data_filename <- "data/titanic.csv"
if (!file.exists(data_filename)) {
  dir.create("data")
  download.file("https://www.openml.org/data/get_csv/16826755/phpMYEkMl", data_filename)
}
```

---

## Python Setup

```{r}
library(reticulate)
py_config()
```

---

## The Python Data Science Toolbox

* **Pandas** (`pd`): the main library for wrangling tabular data in Python. (analogous to *tidyverse*) 
* **NumPy** (`np`): the underlying math library. Gives us `array`s of numbers. Conventionally imported as `np`.
* **scikit-learn**: the main library for machine learning in Python.

```{python}
import pandas as pd
import numpy as np
```

---

class: middle, center

## Pandas

---

### Loading data

Data frames in R are automatically converted into Pandas `DataFrame`s:

```{r}
titanic <- read_csv("data/titanic.csv", na = "?")
```

```{python}
r.titanic.__class__
```

Pandas can read CSV files itself. (CSV is such a quirky data format, so read the docs
for all the parameters you can set.)

```{python}
titanic = pd.read_csv("data/titanic.csv", na_values="?")
```

---

class: center, middle

### Exploring data structure

---

```{python}
titanic.shape
num_people, num_variables = titanic.shape
print(f"{num_people} people, {num_variables} variables about each")
```


```{python}
titanic.head()
```

---

```{python}
titanic.info()
```

---

```{python}
titanic.describe()
```

---

### Tidying data

---

#### Drop unneeded columns

```{python}
titanic2 = titanic.drop(['ticket', 'body'], axis = 1)
```


#### Rename columns

```{python}
titanic3 = titanic2.rename(columns={
  "pclass": "passenger_class",
  "survival": "survived",
  "sibsp": "num_siblings_or_spouses_aboard",
  "parch": "num_parents_or_children_aboard",
  "ticket": "ticket_num",
  "embarked": "embarked_from_port",
  "boat": "lifeboat",
})
```

Note that most (but not all!) Pandas methods make a *new* `DataFrame` (they don't modify the existing one).


---

### Dropping missing data

This dataset has a lot of missing data in some columns. For demonstration
purposes, we'll drop people where this data is missing, without
investigating why. But in general:

**Be careful about dropping missing data if you don't know why it's missing**!

```{python}
titanic4 = titanic3.dropna(subset = ['age', 'fare', 'embarked_from_port'])
```


---

### Querying data

.pull-left[
Each column of a `pd.DataFrame` is a `pd.Series`, which is a NumPy `array` with (optional) labels.

```{python}
titanic4['passenger_class']
```
]

.pull-right[
You can use Boolean operations on a `Series` to get another `Series`:

```{python}
is_first_class = titanic3['passenger_class'] == 1
is_first_class
```

How many rows does this Series have? How many columns?
]


---

### Filtering data

.pull-left[
You can use a Boolean series to query data. This syntax means: *keep only
the rows where the value is `True`:

```{python}
titanic4[is_first_class]
```
]

.pull-right[
You can combine queries using Boolean operations (but they need to be the
element-wise versions--`&`, `|`, and `~` instead of `and`, `or`, and `not`).

```{python}
had_companions = titanic3['num_siblings_or_spouses_aboard'] > 0
titanic4[is_first_class & had_companions]
```
]

---

### Counting

You can get the counts of how many times each item occurs in a `Series`:

```{python}
titanic4['passenger_class'].value_counts()
```

---

### Separating data into features and outcomes

sklearn needs the features to be in a separate data frame from the outcomes,
so we need to split them apart ourselves. If we want to predict survival,
we can create `y` as:

```{python}
y = titanic4['survived'] == 1
```

To create `X`, we can either drop the columns we don't want (see "tidying data" above)
or directly ask for a list of columns we do want:

```{python}
#titanic3.columns
numeric_features = [
  'age', 'num_siblings_or_spouses_aboard', 'num_parents_or_children_aboard',
]
categorical_features = ['passenger_class', 'sex', 'embarked_from_port']

X = titanic4[numeric_features]
```

---

```{python}
X.info()
```



---

class: middle, center

## Scikit-Learn (`sklearn`)

---

The documentation is very well structured:

* the [User Guide](https://scikit-learn.org/stable/user_guide.html) gives narrative
  documentation with background and examples (e.g., [logistic regression](https://scikit-learn.org/stable/modules/linear_model.html#logistic-regression))
* the [API Reference](https://scikit-learn.org/stable/modules/classes.html) gives
  the nitty-gritty details about individual classes and functions
* the [Examples](https://scikit-learn.org/stable/auto_examples/index.html) show
  worked examples of using most components.

It's conventional to import only what you actually need from `sklearn`:

```{python}
from sklearn.linear_model import LogisticRegression
from sklearn.model_selection import train_test_split, cross_validate
from sklearn.compose import make_column_transformer
from sklearn.pipeline import make_pipeline
from sklearn.preprocessing import StandardScaler, OneHotEncoder
from sklearn.metrics import accuracy_score, precision_score, recall_score
```

---

### Classifier API

All classifiers have the same basic interface: construct, `fit`, and `predict`.

```{python}
clf = LogisticRegression()
clf.fit(X, y);
y_pred = clf.predict(X)
```

---

### Metrics

The `sklearn.metrics` module implements a [variety of useful metrics](https://scikit-learn.org/stable/modules/model_evaluation.html#model-evaluation).

```{python}
accuracy_score(y_true = y, y_pred = y_pred)
precision_score(y_true = y, y_pred = y_pred)
recall_score(y_true = y, y_pred = y_pred)
```

([Remember](https://en.wikipedia.org/wiki/Sensitivity_and_specificity#Confusion_matrix):
"recall" = true positive rate = sensitivity; "precision" = positive predictive value)

---

### Train-Test Split


```{python}
np.random.seed(0)
X_train, X_test, y_train, y_test = train_test_split(
  X, y, test_size=0.1)
X_train.shape, y_train.shape
X_test.shape, y_test.shape
```


---

### Cross Validation

```{python}
cv_results = cross_validate(clf, X_train, y_train, cv=5,
  scoring=['accuracy', 'precision', 'recall'])

# Wrap the results in a DataFrame:
cv_results = pd.DataFrame(cv_results).reset_index()
```

We can now access this data in R.

```{r}
py$cv_results
```

---

### Column Transformers

Column transformers let us apply preprocessing steps to subsets of columns.
For example, we'll scale the numeric features:

```{python}
numeric_feature_proc = StandardScaler()
```

and one-hot-encode the categorical features:

```{python}
categorical_feature_proc = OneHotEncoder()
```

And we'll apply each pre-processor to its corresponding columns:

```{python}
preprocessor = make_column_transformer(
  (numeric_feature_proc,     numeric_features),
  (categorical_feature_proc, categorical_features),
  remainder = 'drop')
```

---

### Pipelines

Pipelines put several steps in sequence. Like *workflows* in `tidymodels`, we
can use pipelines to say that the data should be preprocessed before running
the model:

```{python}
clf = make_pipeline(preprocessor, LogisticRegression())
```

Now we can use all of our features!

```{python}
X = titanic4.drop(["survived"], axis = 1)
```

Redo the train-test split:

```{python}
np.random.seed(0)
X_train, X_test, y_train, y_test = train_test_split(
  X, y, test_size=0.1)
```


---

### Pipelines have the model API

```{python}
clf.fit(X, y)
```


---

### CV with pipelines

A pipeline behaves exactly like a classifier (it has `fit` and `predict`),
so we can use exactly the same code to validate it.

```{python}
cv_results = cross_validate(clf, X_train, y_train, cv=5,
  scoring=['accuracy', 'precision', 'recall'])

# Wrap the results in a DataFrame:
cv_results = pd.DataFrame(cv_results).reset_index()
```

We can now access this data in R.

```{r}
py$cv_results
```
