---
title: "Classification"
author: "K Arnold"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: FALSE
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
source("../slides-common.R")
slideSetup(mark_languages = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

* Apply the `tidymodels` pipeline to a classification task
* Identify corresponding elements between R and Python data wrangling and modeling workflows

## Outline

* Problem introduction
  * Data wrangling in R
* Classification workflow
  * Models: decision tree, logistic regression
  * Model outputs: scores and decisions
  * Model metrics: accuracy, sensitivity, specificity
  * Validation
* Python
  * Data wrangling with Pandas
  * scikit-learn model API: construct, fit, predict
  * scikit-learn modeling pipelines


---

## Setup

.pull-left[
```{r}
library(tidyverse)
library(tidymodels)
```
]

.pull-right[

```{r}
library(reticulate)
py_discover_config()
```

```{python}
import pandas as pd
from sklearn.linear_model import LogisticRegression
```
]

---

## Can a blood test diagnose autism?

We'll use an example from <https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005385#pcbi.1005385.s001>

```{r}
data_filename <- "data/autism.csv"
if (!file.exists(data_filename)) {
  dir.create("data")
  download.file("https://doi.org/10.1371/journal.pcbi.1005385.s001", data_filename)
}
```

```{r}
col_names <- names(read_csv(data_filename, n_max = 1, col_types = cols(.default = col_character())))
autism <- read_csv(data_filename, skip = 2, col_names = col_names, col_types = cols(
  .default = col_double(),
  Group = col_character()
))
```

```{python}
autism = r.autism
```

---

We have data about `r nrow(autism)` children:

```{r}
dim(autism)
```

```{python}
autism.shape
```


The data include 3 groups of columns:

1. The outcome (`Group`): ASD (diagnosed with ASD), SIB (sibling not diagnosed with ASD), and NEU (age-matched neurotypical children,
for control)

```{r}
autism %>% group_by(Group) %>% summarize(n = n())
```


```{python}
autism['Group'].value_counts().to_frame()
```


2. Concentrations of various metabolites in a blood sample:

```{r}
autism %>% select(-1, -last_col())
```

```{python}
autism.iloc[:, 1:-1]
```

3. Adaptive behavior score (a measure of life skills)

```{r}
autism %>% 
  ggplot(aes(x = `Vineland ABC`, y = Group)) + geom_boxplot()
```

---

```{r}
autism %>% head()
```


```{python}
autism.info()
```

---

## EDA

```{r fig.width=10, fig.asp=1}
without_behavior_scores <- autism %>% select(-`Vineland ABC`)
without_behavior_scores %>% 
  pivot_longer(-Group, names_to = "Measure") %>% 
  ggplot(aes(x = value, y = Group)) +
  ggridges::geom_density_ridges() +
  facet_wrap(vars(Measure), scales = "free_x") + 
  ggridges::theme_ridges()
```


---

## Can we predict ASD vs non-ASD from metabolites?

Stratify by `group` to ensure each fold has good representation.

```{r}
no_sibs <- without_behavior_scores %>% filter(Group != "SIB")

resamples <- no_sibs %>% vfold_cv(v = 10, strata = Group)
```

```{r}
spec <- workflow() %>% add_recipe(
  recipe(Group ~ ., data = without_behavior_scores)) %>% 
  add_model(logistic_reg() %>% set_engine("glm"))

cv_results <- spec %>% 
  fit_resamples(resamples, metrics = metric_set(yardstick::accuracy, yardstick::sensitivity, yardstick::specificity))
```

```{r}
cv_results %>% 
  collect_metrics(summarize = FALSE) %>% 
  ggplot(aes(x = .estimate, y = .metric)) + geom_boxplot()
```

```{r}
spec <- workflow() %>% add_recipe(
  recipe(Group ~ ., data = no_sibs) %>% step_normalize(all_numeric())) %>% 
  add_model(
    #parsnip::decision_tree(mode = "classification") %>% set_engine("rpart")
    logistic_reg(penalty = 1.0) %>% set_engine("glmnet")
  )

cv_results <- spec %>% 
  fit_resamples(resamples, metrics = metric_set(yardstick::accuracy, yardstick::sensitivity, yardstick::specificity))
```

```{r}
cv_results %>% 
  collect_metrics(summarize = FALSE) %>% 
  ggplot(aes(x = .estimate, y = .metric)) + geom_boxplot()
```

```{r}
model <- spec %>% fit(no_sibs)
```

```{r}
model %>% pull_workflow_fit() %>% pluck('fit') %>% coef(s = .1) %>% as.matrix() %>% as_tibble(rownames = "name") %>% 
  rename(coef = 2) %>% 
  ggplot(aes(x = coef, y = fct_reorder(name, coef, abs))) + geom_col()
```


```{python}
no_sibs = r.no_sibs
y = no_sibs['Group']
X = no_sibs.drop(['Group'], axis=1)
clf = LogisticRegression(max_iter = 1e6)
clf.fit(X, y)
```

```{python}
clf.intercept_
```


```{python}
import io
import pandas as pd
output = io.StringIO()
pd.DataFrame({'text': ["a\"\r\"\n\n\n\nb"]}).to_csv(output, index=False)
pd.read_csv(io.StringIO(output.getvalue()))
print(output.getvalue())
```

