---
title: "Classification"
author: "K Arnold"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: FALSE
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
source("../slides-common.R")
slideSetup(mark_languages = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

* Apply the `tidymodels` pipeline to a classification task
* Identify corresponding elements between R and Python data wrangling and modeling workflows

## Outline

* Problem introduction
  * Data wrangling in R
* Classification workflow
  * Models: decision tree, logistic regression
  * Model outputs: scores and decisions
  * Model metrics: accuracy, sensitivity, specificity
  * Validation
* Python
  * Data wrangling with Pandas
  * scikit-learn model API: construct, fit, predict
  * scikit-learn modeling pipelines


---

## Setup

```{r}
library(tidyverse)
library(tidymodels)
```

```{r utilities}
add_predictions <- function(data, ...) {
  imap_dfr(
    rlang::dots_list(..., .named = TRUE),
    function(model, model_name) {
      model %>%
        predict(data) %>%
        bind_cols(data) %>%
        mutate(model = !!model_name)
    }
  )
}

sweep_model_examples <- function(model, dataset, vars_to_sweep, examples = slice_sample(dataset, n = 10)) {
  X <- map_dfr(vars_to_sweep, function(v) {
    var_to_sweep <- rlang::as_label(v)
    sweep_min <- min(dataset[[var_to_sweep]])
    sweep_max <- max(dataset[[var_to_sweep]])
    expand_grid(
      examples %>% select(-!!var_to_sweep) %>% mutate(.idx = row_number()),
      !!enquo(var_to_sweep) := seq(sweep_min, sweep_max, length.out = 500)) %>% 
      mutate(sweep_var = var_to_sweep, .sweep_val = .data[[var_to_sweep]])
  })
  model %>% 
    predict(X) %>% 
    bind_cols(X)
}
```


---

## Can a blood test diagnose autism?

We'll use an example from <https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005385#pcbi.1005385.s001>

```{r download-data}
data_filename <- "data/autism.csv"
if (!file.exists(data_filename)) {
  dir.create("data")
  download.file("https://doi.org/10.1371/journal.pcbi.1005385.s001", data_filename)
}
```

```{r read-data}
col_names <- names(read_csv(data_filename, n_max = 1, col_types = cols(.default = col_character())))
autism <- read_csv(data_filename, skip = 2, col_names = col_names, col_types = cols(
  .default = col_double(),
  Group = col_character()
)) %>% mutate(
  Group = as_factor(Group)
)
```

---

We have data about `r nrow(autism)` children:

```{r dim-data}
dim(autism)
```


The data include 3 groups of columns:

1. The outcome (`Group`): ASD (diagnosed with ASD), SIB (sibling not diagnosed with ASD), and NEU (age-matched neurotypical children,
for control)

```{r group-counts}
autism %>% group_by(Group) %>% summarize(n = n())
```

2. Concentrations of various metabolites in a blood sample:

```{r show-metabolites}
autism %>% select(-1, -last_col())
```

3. Adaptive behavior score (a measure of life skills)

```{r show-behavior-score}
autism %>% 
  ggplot(aes(x = `Vineland ABC`, y = Group)) + geom_boxplot()
```

---

## EDA

```{r fig.width=10, fig.asp=1}
autism %>%
  select(-`Vineland ABC`) %>% 
  pivot_longer(-Group, names_to = "Measure") %>% 
  ggplot(aes(x = value, y = Group)) +
  ggridges::geom_density_ridges() +
  facet_wrap(vars(Measure), scales = "free_x") + 
  ggridges::theme_ridges()
```


---

## Can we predict ASD vs non-ASD from metabolites?

Stratify by `group` to ensure each fold has good representation.

```{r}
data <- 
  autism %>% 
  select(-`Vineland ABC`) %>% 
  filter(Group != "SIB")
```

```{r}
spec <- workflow() %>% add_recipe(
  recipe(Group ~ ., data = data)) %>% 
  add_model(logistic_reg() %>% set_engine("glm"))
```

```{r}
model <- spec %>% fit(data)
```

---

```{r predicted-probs}
model %>% predict(data, type = "prob")
```
```{r plot-predicted-probs}
model %>%
  predict(data, type = "prob") %>%
  #bind_cols(predict(model, data, type = "class")) %>% 
  bind_cols(data) %>% 
  mutate(idx = row_number()) %>% 
  ggplot(aes(x = idx, y = .pred_ASD, color = Group, shape = Group)) +
  geom_hline(yintercept = .5) +
  geom_point() 
```

```{r}
model %>% 
  predict(data, type = "class") %>% 
  bind_cols(data) %>% 
  yardstick::sensitivity(truth = Group, estimate = .pred_class)
```


---


```{r}
model %>% pull_workflow_fit() %>% pluck('fit') %>% coef(s = .1) %>% as.matrix() %>% as_tibble(rownames = "name") %>% 
  rename(coef = 2) %>% 
  ggplot(aes(x = coef, y = fct_reorder(name, coef, abs))) + geom_col()
```

---


```{r}
resamples <- data %>% vfold_cv(v = 10, strata = Group)
```


```{r}
cv_results <- spec %>% 
  fit_resamples(resamples, metrics = metric_set(yardstick::accuracy, yardstick::sensitivity, yardstick::specificity))
```

```{r}
cv_results %>% 
  collect_metrics(summarize = FALSE) %>% 
  ggplot(aes(x = .estimate, y = .metric)) + geom_boxplot()
```

```{r}
spec <- workflow() %>% add_recipe(
  recipe(Group ~ ., data = data) %>% step_normalize(all_numeric())) %>% 
  add_model(
    #parsnip::decision_tree(mode = "classification") %>% set_engine("rpart")
    logistic_reg(penalty = 0.001) %>% set_engine("glmnet")
  )

cv_results <- spec %>% 
  fit_resamples(resamples, metrics = metric_set(yardstick::accuracy, yardstick::sensitivity, yardstick::specificity))
```

```{r}
cv_results %>% 
  collect_metrics(summarize = FALSE) %>% 
  ggplot(aes(x = .estimate, y = .metric)) + geom_boxplot()
```

---

## Python

```{r}
library(reticulate)
py_discover_config()
```

```{python}
import pandas as pd
from sklearn.linear_model import LogisticRegression
```




```{python}
autism = r.autism
```


```{python}
autism.shape
```


```{python}
autism.info()
```


```{python}
autism['Group'].value_counts().to_frame()
```



```{python}
autism.iloc[:, 1:-1]
```



```{python}
data = r.data
y = data['Group']
X = data.drop(['Group'], axis=1)
clf = LogisticRegression(max_iter = 1e6)
clf.fit(X, y)
```

```{python}
clf.intercept_
```

