---
title: "Quiz 8 Supplement"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(learnr)
library(tidyverse)
library(tidymodels)
# Disable a warning message.
options(dplyr.summarise.inform = FALSE)
theme_set(theme_light())

car_prices <- readxl::read_excel("data/kuiper.xlsx") %>% janitor::clean_names()
fuel_eco <- read_csv("data/guide2005-2004oct15.csv") %>% janitor::clean_names()

set.seed(0)
cars_split <- initial_split(car_prices, prop = 3/4, strata = "make")
cars_train <- training(cars_split)
cars_test <- testing(cars_split)
```

## Visualization

This quiz uses a dataset from [Kuiper 2017](https://www.tandfonline.com/doi/full/10.1080/10691898.2008.11889579)
of car sale prices from Kelly Blue Book for several hundred 2005 used General Motors (GM) cars.
The [link to the data set](http://ww2.amstat.org/publications/jse/v16n3/kuiper.xls) can be found at the bottom of the article,
if you're interested in running this analysis on your own computer.

```{r load-data, eval=FALSE}
car_prices <- readxl::read_excel("data/kuiper.xlsx") %>% 
  janitor::clean_names()
```

Let's take a quick look at the data:

```{r show-data, echo=TRUE}
car_prices %>% head()
car_prices %>% glimpse()
```

### Reproduce a plot

#### Mileage

Write code to reproduce the following visualization, intended to show that
lower-mileage cars tend to be valued more highly.

Tips:

* See the [Visualization Tweaks](https://cs.calvin.edu/courses/data/202/fa20/notes/visualization.html#tweaks)
  section of the course notes for tips on the scales.

```{r mileage-plot, echo=FALSE}
car_prices %>% 
  ggplot(aes(x = mileage, y = price / 1000, color = type, shape = type)) +
  geom_point() +
  scale_x_continuous(labels = scales::comma) +
  labs(x = "Mileage (mi)", y = "Price ($1000)") + 
  theme_light()
```

```{r mileage-plot-reproduce, exercise=TRUE}
car_prices %>% ___
```



Extra credit: can you figure out how to make this plot? Email me your answer.
Tip: two `geom_point` layers, one with `data` set to `car_prices` *without the `make` column*.

```{r mileage-plot-improved, echo=FALSE}
car_prices %>%
  ggplot(aes(x = mileage / 1000, y = price / 1000, color = make, shape = make)) +
  geom_point(data = car_prices %>% select(-"type"), color = "grey", size = .5) + 
  geom_point(size = 1) +
  facet_wrap(vars(type)) + 
  labs(x = "Mileage (1000 mi)", y = "Price ($1000)")
```


### Mileage effect by price

Write code to reproduce the following visualization, intended to show whether
mileage affects price differently for different types of cars.
Consider a car to be "low mileage" if its mileage is less than the median
mileage.


```{r mileage-lo-hi, echo=FALSE}
car_prices %>% 
  #group_by(type) %>% 
  mutate(low_mileage = mileage < median(mileage)) %>% 
  ggplot(aes(y = fct_reorder(type, price), x = price / 1000, color = low_mileage)) +
    geom_boxplot() +
    labs(y = "", x = "Price ($1000)", color = "Low mileage") +
    xlim(c(0, NA))
```

```{r mileage-lo-hi-reproduce, exercise=TRUE}

```

## Data Wrangling

What was the most common car `make` in this data set?

```{r most-common-make, echo=FALSE}
car_prices %>% 
  group_by(make) %>%
  summarize(n = n()) %>% 
  arrange(desc(n))
```

```{r most-common-make-repro, exercise=TRUE}

```

Make a table showing, for each car `type`, what fraction of the listed cars had
cruise control. Note that the `cruise` column contains a 1 if the listed car
has cruise control, 0 if it doesn't. Your table should have 5 rows.
**Sort the table from highest to lowest fraction**.

Which car type **least frequently** had cruise control?

```{r include=FALSE}
car_prices %>%
  group_by(type) %>% 
  summarise(frac_cruise = mean(cruise)) %>% 
  arrange(desc(frac_cruise))
```

```{r cruise-control-rep, exercise=TRUE}

```


Do 2-door cars or 4-door (`doors`) cars have bigger engines (`liter` column) on average in this data set?

```{r include=FALSE}
car_prices %>%
  #distinct(Make, Model, Trim, Doors, Liter) %>% 
  group_by(doors) %>% 
  summarize(mean_engine_size = mean(liter))
```

```{r doors-engine-size, exercise=TRUE}

```

<!--
Which car (`make` and `model`) has the most different trim levels?

```{r}
car_prices %>% 
  group_by(make, model, trim) %>%
  summarize(n = n()) %>% 
  group_by(make, model) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```
```{r}
car_prices %>%
  group_by(make, trim) %>% 
  summarise(n = n()) %>% 
  group_by(make) %>% 
  summarise(n = sum(n))
```

-->

### Joining

For increasingly many consumers, fuel economy is a deciding factor in purchasing
decisions. However, the Kelly Blue Book data set doesn't include any fuel economy
data. Fortunately the US Department of Energy and the EPA put together
<https://www.fueleconomy.org>. That site provides both consumer-facing tools
to compare the monetary and environmental fuel costs of various vehicles.
Fortunately they also release machine-readable [data files](https://www.fueleconomy.gov/feg/download.shtml).
According to the Kuiper 2017 article, all of the vehicles were 2005 model year
and less than a year old, so the 2005 data has been downloaded for you:

```{r load-mpg, eval=FALSE, echo=TRUE}
fuel_eco <- read_csv("data/guide2005-2004oct15.csv") %>% janitor::clean_names()
```

Notable columns are (descriptions are quoted from the EPA's [guide](https://www.fueleconomy.gov/feg/pdfs/guides/FEG2005.pdf)):

* `manufacturer` and `carline_name`: the make and model of the car tested.
* `cty`: *City* represents urban driving, in which a vehicle is started in the morning (after being parked all night) and driven in stop-and-go rush hour traffic
* `hwy`: *Highway* represents a mixture of rural and interstate highway driving in warmed-up vehicles, typical of longer trips in free-flowing traffic.
* `cmb`: *Combined* city and highway MPG estimates ... assume you will drive 55% in the city and 45% on the highway.

Full documentation from the source is available [here](https://www.fueleconomy.gov/feg/epadata/Readme.txt).

We first briefly verify that the *carline name* is sufficiently unique to describe a car; there
are no cars with the same name but different manufacturers:

```{r echo=TRUE}
fuel_eco %>% distinct(carline_name) %>% nrow()
fuel_eco %>% distinct(manufacturer, carline_name) %>% nrow()
```

Good to go. But a brief look at the dataset immediately reveals two potential problems:

```{r echo=TRUE}
fuel_eco %>% select(manufacturer, carline_name, cty, hwy, cmb) %>% head(5)
```

First problem: each car may be reported several times in the EPA dataset, once
for each different engine option or transmission type. We could work out
some of that data from the `trim` column 

Let's look at `car_prices` again for reference:

```{r echo=TRUE}
car_prices %>% distinct(make, model) %>% head(5)
```


Second problem: the EPA uses all-caps for car names but the Kelly Blue Book didn't.
So none of the EPA's `carline_name`s in `fuel_eco` will match the `model`s in `car_prices`.
But we won't let that stop us! Here's a strategy: ignore case. Specifically, 
let's *upper-case* (`str_to_upper()`) the models in `car_prices` and  put the result in a *new column*. 
Then we can use that new column to match up the cars between the two data frames.

1. Use this case-ignoring strategy to **look up the highway fuel economy** for
   all of the cars with matching model names. This strategy should work for
   some but not all of the `r nrow(car_prices)` cars in the Blue Book dataset.
   **How many cars matched**?
2. For the cars in the Blue Book dataset that the case-ignoring strategy successfully matched,
  which `type` of car (convertible, coupe, ...) has the highest median highway MPG?
3. Give an example of a car for which the lookup failed.
4. Name a strategy for fixing the mismatch.

```{r include=FALSE}
car_prices_and_eco <- 
  car_prices %>% 
  mutate(carline_name = str_to_upper(model)) %>% 
  left_join(
    fuel_eco %>% 
      group_by(carline_name) %>% 
      summarize(hwy = mean(hwy)),
    by = "carline_name")

car_prices_and_eco %>% 
  summarize(sum(!is.na(hwy)))

car_prices_and_eco %>% 
  drop_na(hwy) %>% 
  ggplot(aes(x = hwy, y = price, color = type)) +
  geom_point() +
  geom_smooth(method = "lm")

car_prices_and_eco %>% 
  drop_na(hwy) %>% 
  group_by(type) %>% 
  summarize(median_hwy = median(hwy))

car_prices_and_eco %>% 
  drop_na(hwy) %>% 
  distinct(type, model, hwy) %>% 
  group_by(type) %>% 
  summarize(median_hwy = median(hwy))

car_prices_and_eco %>% 
  filter(is.na(hwy)) %>% 
  distinct(make, model)
```


## Modeling

Let's try to make some models to predict `price`.

We'll hold out a testing set so we can validate our model.

```{r train-test-split, echo=TRUE}
set.seed(0)
cars_split <- initial_split(car_prices, prop = 3/4, strata = "make")
cars_train <- training(cars_split)
cars_test <- testing(cars_split)
```


First, create a `tidymodels` model called `model` that predicts `price` from `mileage` 
using linear regression on the training set. (You do not need a recipe here.)

Then, write code to recreate the following plot, which shows how the 
*residuals* (actual minus predicted) depend on the car type.
Plots like this can be useful for showing whether it's helpful to add additional
features to the predictive model.

(Note: because of the limitations of this platform, you'll need to do both the
model fitting and the graph creation in the same cell.)

```{r residual-by-type-plot, eval=TRUE}
model <- 
  parsnip::linear_reg() %>% 
  set_engine('lm') %>% 
  fit(price ~ mileage, data = cars_train)

model %>% 
  predict(cars_train) %>% 
  bind_cols(cars_train) %>% 
  mutate(residual = price - .pred) %>% 
  ggplot(aes(y = fct_reorder(type, residual), x = residual)) +
    geom_vline(xintercept = 0) +
    geom_boxplot() +
    labs(y = "")
```

```{r residual-by-type-repro, exercise=TRUE}

```

