---
title: "Quiz 8 Supplement"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(learnr)
library(tidyverse)
library(tidymodels)
# Disable a warning message.
options(dplyr.summarise.inform = FALSE)

car_prices <- readxl::read_excel("data/kuiper.xlsx")
```

## Visualization

This quiz uses a dataset from [Kuiper 2017](https://www.tandfonline.com/doi/full/10.1080/10691898.2008.11889579)
of car sale prices from Kelly Blue Book for several hundred 2005 used General Motors (GM) cars.

The [link to the data set](http://ww2.amstat.org/publications/jse/v16n3/kuiper.xls) can be found at the bottom of the article,
if you're interested in running this analysis on your own computer.

```{r load-data, eval=FALSE}
car_prices <- readxl::read_excel("data/kuiper.xlsx")
```


```{r show-data}
car_prices %>% head()
car_prices %>% glimpse()
```


### Reproduce a plot

Write code to reproduce the following visualization, intended to show that
lower-mileage cars tend to be valued more highly.

Tips:

* See the [Visualization Tweaks](https://cs.calvin.edu/courses/data/202/fa20/notes/visualization.html#tweaks)
  section of the course notes for tips on the scales.
* Add in `theme_light()` to get the plot styling (optional).

```{r mileage-plot, echo=FALSE}
car_prices %>% 
  ggplot(aes(x = Mileage, y = Price / 1000, color = Make, shape = Make)) +
  geom_point() +
  scale_x_continuous(labels = scales::comma) +
  labs(x = "Mileage (mi)", y = "Price ($1000)") + 
  theme_light()
```

```{r mileage-plot-reproduce, exercise=TRUE}
car_prices %>% ___
```



Extra credit: can you figure out how to make this plot? Email me your answer.
Tip: two `geom_point` layers, one with `data` set to `car_prices` *without the `Make` column*.

```{r mileage-plot-improved, echo=FALSE}
car_prices %>%
  ggplot(aes(x = Mileage, y = Price)) +
  geom_point(data = car_prices %>% select(-Make), color = "grey", size = .5) + 
  geom_point(size = 1) +
  facet_wrap(vars(Make)) + 
  theme_minimal()
```

<!--

Write code to reproduce the following visualization, intended to show how


```{r mileage-plot, echo=FALSE}
car_prices %>% 
  group_by(Make) %>% 
  mutate(low_mileage = Mileage < median(Mileage)) %>% 
  ggplot(aes(y = fct_reorder(Make, Price), x = Price, color = low_mileage)) +
    geom_boxplot()

car_prices %>% 
  mutate(low_mileage = Mileage < median(Mileage)) %>% 
  ggplot(aes(y = fct_reorder(Make, Price), x = Price, color = low_mileage)) +
    geom_boxplot()
```

-->

## Data Wrangling

What was the most common car `Make` in this data set?

```{r echo=FALSE}
car_prices %>% 
  group_by(Make, Model, Trim) %>%
  summarize(n = n()) %>% 
  arrange(desc(n))
```

<!--
Which car (`Make` and `Model`) has the most different trim levels?

```{r}
car_prices %>% 
  group_by(Make, Model, Trim) %>%
  summarize(n = n()) %>% 
  group_by(Make, Model) %>% 
  summarize(n = n()) %>% 
  arrange(n)
```

-->

Do 2-door cars or 4-door (`Doors`) cars have bigger engines (`Liter` column) on average in this data set?

```{r}
car_prices %>%
  #distinct(Make, Model, Trim, Doors, Liter) %>% 
  group_by(Doors) %>% 
  summarize(size = mean(Liter))
```

## Modeling

Let's try to make some models to predict `Price`.

We'll hold out a testing set so we can validate our model.

```{r train-test-split}
set.seed(0)
cars_split <- initial_split(car_prices, prop = 3/4, strata = "Make")
cars_train <- training(cars_split)
cars_test <- testing(cars_split)
```


Create a `tidymodels` model called `model` that predicts `Price` from `Mileage` 
using linear regression on the training set.

```{r}
trained_model <- 
  parsnip::linear_reg() %>% 
  set_engine('lm') %>% 
  fit(Price ~ Mileage, data = cars_train)
```

