---
title: "Feature Engineering and Review"
author: "K Arnold"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: FALSE
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
source("../slides-common.R")
slideSetup()
knitr::opts_chunk$set(echo = TRUE)
```

## Logistics

* Project: outline is posted
* Midterm Quiz posted


Plan:

* TODAY: review
* Wed: decision trees
* Fri: lab about **overfitting**

---

## General Hints

* Visualization:
  * What *glyph* represents each *observation*? 
  * What *attributes*/aesthetics does that glyph have? (x, y, width, color, ...)
  * What *controls* each aesthetic? ("each party has a *y* position", ...)
* Wrangling
  * What does the *input* look like? (Translate the first row into a data sentence in English.)
  * What does the *output* need to look like? (Again, write a sentence.)
  * What *sequence of steps* needs to happen? (e.g., `filter-group_by-summarize-arrange`)
* Modeling
  * What *quantity* are you trying to predict?
  * What *error measure* will tell you the prediction is good / bad?
  * What *features* can help you make that prediction?

---

## Q&A

> Recipes vs Data Wrangling Pipelines?

* recipe (from `recipes` package) = data wrangling pipeline
  * ... that can be easily applied to new data (e.g., *test set*)
  * ... that can have learnable state (like ranges of data values)
  
> Why did the prediction on the example test set house come out the same when we re-scaled the data?

* We'd applied *exactly the same* transformation to the example house
  as we did to the training data.
* Linear regression is *linear*: it doesn't care what units the data is in.
* (So the specific *range* didn't matter.)

> Are we still going to work in cohorts/teams?

Friday's lab, hopefully.

---

```{r}
library(tidymodels)
data(ames, package = "modeldata")
ames <- ames %>% 
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal") %>% 
  mutate(across(where(is.integer), as.double))
```

```{r train-test-split}
set.seed(10) # Seed the random number generator
ames_split <- initial_split(ames, prop = 2/3) # Split our data randomly
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

We'll use one example home from the test set.

```{r example-home}
example_home <- ames_test %>% slice(1)
example_home %>% select(Gr_Liv_Area, Sale_Price)
```

---

class: center, middle

# Recipes

---

```{r}
ames_recipe <- 
  recipe(
    Sale_Price ~ Latitude + Longitude + Neighborhood + Year_Sold + Gr_Liv_Area,
    data = ames_train
  ) %>% 
  step_other(Neighborhood) %>% 
  step_dummy(all_nominal()) %>% 
  #step_interact(~ starts_with("Neighborhood_") : Year_Sold) %>% 
  step_center(all_predictors()) %>% 
  step_scale(all_predictors()) %>% 
  step_log(Sale_Price, base = 10) %>% 
  prep(trainig = ames_train, retain = TRUE)
ames_recipe %>% summary()
```

---

```{r}
ames_recipe %>% bake(new_data = ames_train)
```

---

## What computations can a linear model do?

* *Add* up terms.
* Each term: *multiply* a number by a constant.

.pull-left[
```{r coefs-from-fitted-model}
intercept <- 15793
coef_living_area <- 108
```


```{r}
intercept + coef_living_area * 1610
```
]

.pull-right[
```{r price-vs-liv-area, out.width="100%"}
ggplot(ames, aes(x = Gr_Liv_Area, y = Sale_Price)) +
  geom_point() +
  geom_hline(yintercept = example_home$Sale_Price, color = "red") +
  geom_vline(xintercept = example_home$Gr_Liv_Area, color = "red") +
  geom_point(data = example_home, color = 'red', size = 5) +
  geom_function(fun = function(x) intercept + coef_living_area * x, color = "blue")
```
]

---

## Do remodeled homes sell for more?

`Year Remod/Add`: Remodel date *(same as construction date if no remodeling or additions)*
(from dataset documentation)

```{r remodeled, out.width = "70%"}
ames %>% 
  mutate(remodeled = Year_Remod_Add != Year_Built) %>% #<< 
  ggplot(aes(x = Gr_Liv_Area, y = Sale_Price, color = remodeled)) +
  geom_point() +
  geom_smooth(method = "lm")
```


---

## Conditional Logic: Simple Conditions

How could a *linear model* treat remodeled homes differently from non-remodeled?

```
if remodeled:
  Sale_Price = intercept_remodeled + coef_sqft * Gr_Liv_Arera
else:
  Sale_Price = intercept_other + coef_sqft * Gr_Liv_Arera
```

--

### Solution: "**dummy encoding**"

```
Sale_Price = 
   intercept_other 
   + coef_remodeled * (1 if remodeled)
   + coef_sqft      * Gr_Liv_Area
```

---

```{r}
ames_train_2 <- ames_train %>% 
  mutate(remodeled = case_when(
    Year_Built == Year_Remod_Add ~ "no",
    TRUE                         ~ "yes") %>%
    as_factor()
  )
```

.small[

```{r}
ames_recipe_3 <- 
  recipe(Sale_Price ~ Gr_Liv_Area + remodeled, data = ames_train_2) %>% 
  step_dummy(remodeled) %>% #<<
  step_range(all_numeric(), -all_outcomes(), min = 0, max = 1) %>%
  prep()
ames_recipe_3 %>% bake(new_data = ames_train_2) %>% head(10) %>% knitr::kable(format = "html")
```
]
