---
title: "Intro"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
library(reticulate)
reticulate::py_run_string(r"--(
import pandas as pd
import plotly.express as px
health_and_wealth_2007 = px.data.gapminder(year=2007)
)--")
```


## Warmup

### Running Python Code

Write the Python code required to add two plus two:

```{python two-plus-two, exercise=TRUE}

```

### Dataset info

Run each of these exercises to explore different ways of seeing what's in a data frame.

Write a comment line (`# `) above each one to document what you observe that it does.

```{python df-info-1, exercise=TRUE, exercise.lines = 2}
# This shows the first and last rows in the data frame.
health_and_wealth
```

```{python df-info-2, exercise=TRUE, exercise.lines = 2}
# This ...
health_and_wealth.info()
```

```{python df-info-3, exercise=TRUE, exercise.lines = 2}
# ...
health_and_wealth.head()
```

```{python df-info-4, exercise=TRUE, exercise.lines = 2}
# ...
health_and_wealth.shape[0]
```



### Making a Plot

Use color to show which continent each country is in. To do this, add a mapping from `color` to `continent`.

```{python plot-gdp, exercise=TRUE, exercise.lines = 5}
health_and_wealth = px.data.gapminder(year=2007)
health_and_wealth.info()
px.scatter(health_and_wealth, x="gdpPercap", y="lifeExp")
```

```{python plot-gdp-solution}
health_and_wealth = px.data.gapminder(year=2007)
health_and_wealth.info()
px.scatter(health_and_wealth, x="gdpPercap", y="lifeExp", color="continent")
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaluated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

