---
title: "Lab 09: Over-fitting"
output: 
  tufte::tufte_html:
    css: ../lab.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
options(scipen = 5) # encourage metrics to print in fixed-point notation
```

## Goals

* Adjust parameters to obtain a model that fits *too* well
* Practice *model validation* on held-out data
* Practice more visualization (as always!)

## Getting started

Find your repo in [our GitHub organization](https://github.com/Calvin-DS202-FA20/)
and clone it as usual.

### Optional, experimental workflow

The RStudio server lets multiple people edit a single project at the same time!
This can be confusing since it's different from collaborating using GitHub, so
try it only if all of your team is game to try:

* The person whose name appears first on the Attendance sheet (or someone else if needed)
  clones the team's repo on the RStudio Server.
* They Share the project with the other team members
* They copy the URL and paste it into the attendance sheet
* Others go to that URL.

## Data

We'll be using the Ames home sales dataset, as we have been using this past week.
Again, see [Data dictionary](http://jse.amstat.org/v19n3/decock/DataDocumentation.txt).
And again, we use a subset of the original data as suggested by the
[original paper](http://jse.amstat.org/v19n3/decock.pdf).

We'll also scale the sale price to be in units of $1000, to keep the numbers more manageable.

```{r load-and-subset-data}
data(ames, package = "modeldata")
ames_all <- ames %>% 
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal") %>% 
  mutate(across(where(is.integer), as.double)) %>% 
  mutate(Sale_Price = Sale_Price / 1000)
rm(ames)
```

We'll use the same train-test split that we have been using:

```{r train-test-split}
set.seed(0) # Seed the random number generator
ames_split <- initial_split(ames_all, prop = 2/3) # Split our data randomly
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

**Note**: unless explicitly mentioned below, use the *training set* for plots
and models.

## Decision Trees

1. Make a plot showing how `Sale_Price` relates to latitude and longitude (make sure you get them the right way!).

```{r include=FALSE}
library(ggrepel)
ggplot(ames_train, aes(x = Longitude, y = Latitude, color = fct_lump_n(Neighborhood, 5))) +
  geom_point() +
  geom_text_repel(
    data = ames_train %>% group_by(Neighborhood) %>% summarize(across(c(Latitude, Longitude), median)),
    aes(label = Neighborhood), color = "black") +
  theme(legend.position = "bottom")
```

```{r}
ggplot(ames_train, aes(x = Longitude, y = Latitude, color = Sale_Price)) +
  geom_point(size = 1) +
  coord_equal()
```



1. Fit a decision tree to predict `Sale_Price` from `Latitude` and `Longitude`.
Use `cost_complexity` = 0.01, `min_n` = 20, and `tree_depth` = 30 (the defaults
for the `rpart` engine).


```{r}
decision_tree_fit <- 
  decision_tree(mode = "regression") %>% #, cost_complexity = 1e-6, min_n = 2, tree_depth = 30) %>%
  set_engine("rpart") %>% 
  fit(Sale_Price ~ Latitude + Longitude, data = ames_train)
```

You can use the following code to visualize the tree:

```{r}
decision_tree_fit$fit %>% rpart.plot::rpart.plot(roundint = FALSE)
```


```{r}
seq_matching_range <- function(x, length.out = 500) {
  seq(min(x), max(x), length.out = length.out)
}

augment_predictions <- function(data, model, variable_name = ".pred") {
  model %>%
    predict(data) %>%
    rename(!!enquo(variable_name) := .pred) %>%
    bind_cols(data)
}


# get_raster <- function(data, model, x_var, y_var, pred_var) {
#   x_var <- enquo(x_var)
#   y_var <- enquo(y_var)
#   pred_var <- enquo(pred_var)
#   example_data <- expand_grid(
#     !!x_var = seq_matching_range(data[[x_var]]),
#     !!y_var = seq_matching_range(data[[y_var]])
#   ) %>%
#     augment_predictions(model)
#   
#   ggplot(data, aes(x = !!x_var, y = !!y_var)) +
#     geom_raster(
#       data = example_data,
#       mapping = aes(fill = .pred)) +
#     geom_point(size = .5)
# }
```


```{r}
lat_long_grid <- expand_grid(
  Latitude = seq_matching_range(ames_train$Latitude),
  Longitude = seq_matching_range(ames_train$Longitude),
  Gr_Liv_Area = median(ames_train$Gr_Liv_Area),
  Total_Bsmt_SF = median(ames_train$Total_Bsmt_SF),
  Bldg_Type = "OneFam"
)

ggplot(ames_train, aes(x = Longitude, y = Latitude)) +
  geom_raster(
    data = augment_predictions(lat_long_grid, decision_tree_fit),
    mapping = aes(fill = .pred)) +
  geom_point(aes(color = Sale_Price), size = .5)
```

```{r}
ames_train %>%
  augment_predictions(decision_tree_fit) %>% 
  mutate(resid = Sale_Price - .pred) %>% 
  ggplot(aes(x = resid)) + geom_histogram(binwidth = 10)
```


```{r}
ames_train %>% 
  augment_predictions(decision_tree_fit) %>% 
  mae(truth = Sale_Price, estimate = .pred)
```


2. Now let's allow the tree to grow: set `cost_complexity` to 0.001.

```{r}
decision_tree_fit2 <- 
  decision_tree(mode = "regression", cost_complexity = 1e-3) %>% # , min_n = 2
  set_engine("rpart") %>% 
  fit(Sale_Price ~ Latitude + Longitude + Gr_Liv_Area + Bldg_Type + Total_Bsmt_SF, data = ames_train)
```


You can use the following code to visualize the tree:

```{r}
#decision_tree_fit2$fit %>% rpart.plot::rpart.plot(roundint = FALSE)
```

```{r}
ggplot(ames_train, aes(x = Longitude, y = Latitude)) +
  geom_raster(
    data = augment_predictions(lat_long_grid, decision_tree_fit2),
    mapping = aes(fill = .pred)) +
  geom_point(aes(color = Sale_Price), size = .5)
```

```{r}
ames_train %>%
  augment_predictions(decision_tree_fit2) %>% 
  mutate(resid = Sale_Price - .pred) %>% 
  ggplot(aes(x = resid)) + geom_histogram(binwidth = 10)
```

Overall, how did the residuals change?


3. Finally, try `cost_complexity = 1e-6` and `min_n = 2`. Since this will make a very large tree,
don't try to run `rpart.plot`.

```{r}
decision_tree_fit3 <- 
  decision_tree(mode = "regression", cost_complexity = 1e-6, min_n = 2) %>% 
  set_engine("rpart") %>% 
  fit(Sale_Price ~ Latitude + Longitude + Gr_Liv_Area + Bldg_Type + Total_Bsmt_SF, data = ames_train)
```

```{r}
ggplot(ames_train, aes(x = Longitude, y = Latitude)) +
  geom_raster(
    data = augment_predictions(lat_long_grid, decision_tree_fit3),
    mapping = aes(fill = .pred)) +
  geom_point(aes(color = Sale_Price), size = .5)
```

```{r}
ames_train %>%
  augment_predictions(decision_tree_fit3) %>% 
  mutate(resid = Sale_Price - .pred) %>% 
  ggplot(aes(x = resid)) + geom_histogram(binwidth = 1)
```

4. Now, evaluate the three models on the test set:


```{r}
expand_grid(
  model = c("decision_tree_fit", "decision_tree_fit2", "decision_tree_fit3"),
  data = c("ames_train", "ames_test")
) %>% pmap_dfr(
  function(model, data) {
    get(data) %>% 
      augment_predictions(get(model)) %>% 
      #mae(truth = Sale_Price, estimate = .pred) %>% 
      #select(mae = .estimate) %>% 
      mutate(resid = Sale_Price - .pred) %>% 
      mutate(model = !!model, data = !!data)
  }) %>% 
#  ggplot(aes(y = model, x = mae)) + geom_col() + facet_wrap(vars(data))
  ggplot(aes(x = resid, y = after_stat(density), color = model)) +
  geom_freqpoly() + facet_wrap(vars(data))
```

