---
title: "Lab 09: Over-fitting"
output: 
  tufte::tufte_html:
    css: ../lab.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
options(scipen = 5) # encourage metrics to print in fixed-point notation
```

## Goals

* Adjust parameters to obtain a model that fits *too* well
* Practice *model validation* on held-out data
* Practice more visualization (as always!)

## Getting started

Find your repo in [our GitHub organization](https://github.com/Calvin-DS202-FA20/)
and clone it as usual.

### Optional, experimental workflow

The RStudio server lets multiple people edit a single project at the same time!
This can be confusing since it's different from collaborating using GitHub, so
try it only if all of your team is game to try:

* The person whose name appears first on the Attendance sheet (or someone else if needed)
  clones the team's repo on the RStudio Server.
* They Share the project with the other team members
* They copy the URL and paste it into the attendance sheet
* Others go to that URL.

## Data

We'll be using the Ames home sales dataset, as we have been using this past week.
Again, see [Data dictionary](http://jse.amstat.org/v19n3/decock/DataDocumentation.txt).
And again, we use a subset of the original data as suggested by the
[original paper](http://jse.amstat.org/v19n3/decock.pdf).

```{r load-and-subset-data}
data(ames, package = "modeldata")
ames_all <- ames %>% 
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal") %>% 
  mutate(across(where(is.integer), as.double))
rm(ames)
```

We'll use the same train-test split that we have been using:

```{r train-test-split}
set.seed(10) # Seed the random number generator
ames_split <- initial_split(ames_all, prop = 2/3) # Split our data randomly
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

**Note**: unless explicitly mentioned below, use the *training set* for plots
and models.

## Decision Trees

1. Make a plot showing how `Sale_Price` relates to latitude and longitude (make sure you get them the right way!).

```{r include=FALSE}
library(ggrepel)
ggplot(ames_train, aes(x = Longitude, y = Latitude, color = fct_lump_n(Neighborhood, 5))) +
  geom_point() +
  geom_text_repel(
    data = ames_train %>% group_by(Neighborhood) %>% summarize(across(c(Latitude, Longitude), median)),
    aes(label = Neighborhood), color = "black") +
  theme(legend.position = "bottom")
```

```{r}
ggplot(ames_train, aes(x = Longitude, y = Latitude, color = Sale_Price)) +
  geom_point(size = 1) +
  coord_equal()
```



1. Fit a decision tree to predict `Sale_Price` from `Latitude` and `Longitude`.
Use `cost_complexity` = 0.01, `min_n` = 20, and `tree_depth` = 30 (the defaults
for the `rpart` engine).


```{r}
decision_tree_fit <- 
  decision_tree(mode = "regression") %>% #, cost_complexity = 1e-6, min_n = 2, tree_depth = 30) %>%
  set_engine("rpart") %>% 
  fit(Sale_Price ~ Latitude + Longitude, data = ames_train)
```


```{r}
decision_tree_fit$fit %>% rpart.plot::rpart.plot(roundint = FALSE)
```


```{r}
seq_matching_range <- function(x, length.out = 500) {
  seq(min(x), max(x), length.out = length.out)
}

example_data <- expand_grid(
  Latitude = seq_matching_range(ames_train$Latitude),
  Longitude = seq_matching_range(ames_train$Longitude)
)

example_data <- decision_tree_fit %>% 
  predict(example_data) %>% 
  rename(Sale_Price = .pred) %>% 
  bind_cols(example_data)

ggplot(ames_train, aes(x = Longitude, y = Latitude)) +
  geom_raster(data = example_data, mapping = aes(fill = Sale_Price)) +
  geom_point(size = .5)
```


