---
title: "Lab 11: Classification"
output: 
  tufte::tufte_html:
    css: ../lab.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(reticulate)
theme_set(theme_bw())
options(scipen = 5) # encourage metrics to print in fixed-point notation
options(dplyr.summarise.inform = FALSE) # silence a warning message
```

## Goals

Learning objectives:

* Basic EDA in Pandas
* Fit a classification model in Scikit-Learn (sklearn)


## Acquiring Data

```{r download-data}
data_filename <- "data/autism.csv"
if (!file.exists(data_filename)) {
  dir.create("data")
  download.file("https://doi.org/10.1371/journal.pcbi.1005385.s001", data_filename)
}
```

```{r read-data}
col_names <- names(read_csv(data_filename, n_max = 1, col_types = cols(.default = col_character())))
autism <- read_csv(data_filename, skip = 2, col_names = col_names, col_types = cols(
  .default = col_double(),
  Group = col_character()
))
```

```{python}
autism = r.autism
```

```{r include=FALSE}
py_repr <- import_builtins()$repr
py_eval_repr <- function(x, quote_code = TRUE) { 
  res <- py_repr(py_eval(x, convert = FALSE))
  if (quote_code) {
    paste0('`', res, '`')
  } else {
    res
  }
}
```

```{python}
import numpy as np
import pandas as pd
from sklearn.linear_model import LogisticRegression
from sklearn.model_selection import train_test_split, cross_validate
from sklearn.compose import make_column_transformer
from sklearn.pipeline import make_pipeline
from sklearn.preprocessing import StandardScaler
from sklearn.metrics import accuracy_score
```


## EDA

1. How many children are represented in this dataset? How many variables are there?

```{python}
num_children, num_variables = autism.shape
print(f"{num_children} children, {num_variables} variables")
```

2. What two variables are unlike the others? What is different about them?

```{python}
autism.info()
```

3. How many children do we have in each `Group`?

```{python}
autism['Group'].value_counts()
```

4. Create a new `DataFrame` called `asd_vs_neu` that has no `SIB` children.

```{python}
asd_vs_neu = autism.query("Group != 'SIB'")
asd_vs_neu = autism[autism['Group'] != "SIB"]
```

5. Verify that the new `DataFrame` has no `SIB` children but the same number of
children in the other groups.

```{python}
asd_vs_neu['Group'].value_counts()
```

`sklearn` has some conventions for what variables are named:

* `X` designates the input data, one row per observation, one column per feature.
  It is uppercase because it's a matrix.
* `y` designates the prediction targets, as a one-dimensional *vector* (and thus lowercase).

6. Create a variable called `y` (designating the *target* of the prediction)
  consisting of the data in the `Group` column of `asd_vs_neu`.

```{python}
y = asd_vs_neu['Group']
```

7. Create a variable called `X` (designating the input data) consisting of 
  all of the columns of `asd_vs_neu` *except for* `Group` and `Vineland ABC`.

```{python}
X = asd_vs_neu.drop(["Group", "Vineland ABC"], axis = 1)
```

8. Check that the `shape` attribute of `X` is `r py_eval_repr('X.shape')`.

```{python}
X.shape
```

Let's hold out a test set of 10% of the children.
First, set the random seed.

```{python}
np.random.seed(0)
X_train, X_test, y_train, y_test = train_test_split(
  X, y, test_size=0.1)
```


Now let's create a `LogisticRegression` object called `clf`, with the
regularization parameter `C` set to 0.001.

```{python}
clf = LogisticRegression(C = 0.001)
```

9. Fit `clf` so that it tries to predict `y_train` from `X_train`.

```{python}
clf.fit(X_train, y_train)
```

10. Construct the model's predictions on the test set. Call it `y_test_pred`.

```{python}
y_test_pred = clf.predict(X_test)
```

11. Evaluate the model's *accuracy* on the test set.

```{python}
accuracy_score(y_true = y_test, y_pred = y_test_pred)
```

12. If, instead, we had wanted to cross-validate:

```{python}
cv_results = cross_validate(clf, X_train, y_train, cv=5,
  scoring=['accuracy'],
  return_train_score=True)
# Wrap the results in a DataFrame:
cv_results = pd.DataFrame(cv_results)
```


```{python}
cv_results['test_accuracy']
```

We can now access this data in R.

```{r}
py$cv_results
```

