---
title: "Lab 04 - Covid Wrangling"
output: 
  tufte::tufte_html:
    css: ../lab.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Covid-19 case counts frequently make headlines these days, but the actual data
is notoriously difficult to work with. The main problems are statistical: testing
is extremely biased, so most case counts are substantial *under*estimates.
But even if we assume that reported counts are somehow useful, there are
problems that come up in analyzing and modeling them that are well suited
to the tools that we are studying this week.

In this lab we will work with Covid-19 case count data from one of the most
prominent sources, the Systems group at Johns Hopkins ([CSEE](https://systems.jhu.edu/)).
In the process of making a few simple plots with this data, we will face and
solve several important data wrangling challenges.

The learning goals of this lab are:

- Recognize some of the issues that come up in real-world data and apply common
  approaches to resolving them
- Manipulate and transform data to prepare it for visualization.
- Start to work together in teams.

## Getting started

Go to [our GitHub org](https://github.com/Calvin-DS202-FA20/) and find your `hw03` repo.

(The "GitHub Classroom" approach has stopped working for us.)

As usual, clone the repo and try to knit the template `Rmd`.


```{r}
library(tidyverse)
library(lubridate)
# Disable a warning message.
options(dplyr.summarise.inform = FALSE)
```


