---
title: "Q&A Misc"
author: "K Arnold"
date: "9/21/2021"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

I'm going to set some options here so these Q&As look like what you'd type at the R console. [docs](https://bookdown.org/yihui/rmarkdown-cookbook/opts-comment.html)

```{r}
knitr::opts_chunk$set(comment = "", prompt = TRUE)
```

## R / RStudio stuff

### What does the `[1]` mean?

R is unusual compared to many other programming languages: most values are actually length-1 lists:

```{r}
1
```

When we have a big list, R spreads it out over multiple lines:

```{r}
numbers <- 1:50
numbers
```

Notice that the number in brackets is the index of the first value on that line:

```{r}
numbers[26]
```

If you don't want that, you can use `cat`:

```{r}
cat(numbers)
```

## What programming language is RStudio?

Natively, R. But you can actually run Python code too with the help of a package called `reticulate`; you just put `python` instead of `r` at the top of a code chunk. See the [cheat sheet](https://raw.githubusercontent.com/rstudio/cheatsheets/master/reticulate.pdf)

## Does saving things on GitHub also save it to your computer?

No, you need to Pull to get any changes from GitHub. I recommend doing all of your edits on RStudio and pushing to GitHub.

## Git / GitHub stuff

### Can you trace the history back to the original commit?

Yes! You can click the History button in RStudio, or go to a GitHub commits page ([example](https://github.com/tidyverse/ggplot2/commits/master)) and keep clicking Older.

