---
title: "Tips for working with SPSS data"
author: "K Arnold"
date: "10/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Pew and other sources release data in a file format used by SPSS, a commercial
statistical analysis tool. Fortunately it's straightforwaard to read this data
in R, using the [`haven`](https://haven.tidyverse.org/) package.

I'll show an example with the American Trends Panel. 

```{r}
atp_w34 <- haven::read_sav("data/W34_Apr18/ATP W34.sav")
```

The easiest way to look at this data is to click on it in the "Environment" panel,
or run `View(atp_w34)` on the Console. (Remember not to leave a `View` call in
an Rmd when you Knit.)

You'll see that each column has a label. It might be hard to read all of them, so
here's a bit of magic code to make a table of just the column labels:

```{r}
getColumnLabels <- function(df) {
  tibble(
    name = names(df),
    label = map_chr(names(df), ~ attr(df[[.]], "label"))
  )
}
getColumnLabels(atp_w34)
```

Many of the columns are actually factors in disguise. To decode their labels,
call `as_factor`. For example, to get party affiliations and leanings from
the ATP data, we can do:

```{r}
atp_w34 %>% 
  mutate(
    party = as_factor(F_PARTY_FINAL),
    party_lean = as_factor(F_PARTYLN_FINAL)) %>% 
  select(party, party_lean)
```
