---
title: "Assignment Management"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# remotes::install_github("rundel/ghclass")
library(ghclass)
library(tidyverse)
library(glue)
library(reticulate)
```

```{r}
org_outside_collaborators <- function(org) {
  ghclass:::ghclass_api_v3_req(endpoint = "GET /orgs/:org/outside_collaborators", org = org) %>%
    purrr::map_chr("login")
}
```


```{r}
ghorg <- "Calvin-DS202-FA20"
```

```{r}
roster <- read_csv("../roster.csv", col_types = cols(
  .default = col_character()
)) %>% janitor::clean_names()
nrow(roster)
stopifnot(
  0 == length(org_members(ghorg) %>% setdiff(org_admins(ghorg)) %>% setdiff(roster$github))
)
```

# Set up assignment templates

```{r}
org_sitrep(ghorg)
```

# Assign Teams


```{r}
fix_roster <- function(roster, assignment, seed) {
  if (assignment != "lab04") {
    roster %>% mutate(
      cohort = case_when(
        str_detect(preferred_email, "bdr22") ~ "H",
        str_detect(preferred_email, "wp22") ~ "B",
        TRUE ~ cohort
      )
    )
  } else {
    roster
  }
}
fix_roster(roster, "lab05", 1) %>% filter(cohort %in% c("B", "H"))
```


```{r assign-teams}
team_seeds <- tribble(
  ~assignment, ~seed,
  "lab04", 0,
  "lab05", 1
)

roster_withteams <- team_seeds %>% 
  pmap_dfr(function(assignment, seed) {
    roster %>%
      fix_roster(assignment, seed) %>% 
    { withr::with_seed(seed, slice_sample(., prop=1)) } %>% 
    group_by(cohort) %>% 
    mutate(
      team = glue("{assignment}-{cohort}{if_else((row_number() - 1) / n() < .5, 1, 2)}")
    ) %>%
    ungroup() %>% 
    mutate(assignment, .before = everything())
  })
roster_withteams

team_sizes <- roster_withteams %>% 
  group_by(assignment, team) %>% 
  mutate(team_size = n()) %>% 
  filter(str_detect(student_name, "Danish") | (assignment != "lab04" & str_detect(student_name, "Brad")))
if (any(team_sizes$team_size != 3)) {
  stop("Danish and Brad should always be on a team of 3.")
}
```

Run this to copy team membership to clipboard. Paste into Excel and copy back out again for easy formatting as HTML table.

```{r copy-team-membership, eval=FALSE}
copy_team_membership <- function(assignment) {
  roster_withteams %>%
    filter(assignment == !!assignment) %>% 
    select(Name = preferred_full_name, Cohort = cohort, Team = team) %>%
    arrange(Team) %>% 
    clipr::write_clip()
}
copy_team_membership("lab05")
```


```{r team-member-count}
roster_withteams %>% filter(assignment == "lab05") %>% count(team)
```

# Create Assignment

```{r}
createAssignment <- function(assignmentTemplatePath, use_template_repo = TRUE) {
  assignmentName <- stringr::str_match(assignmentTemplatePath, r"{\w+/([a-z0-9]+)}")[,2]
  message(glue("Creating assignment {assignmentName}"))
  repoName <- paste0(assignmentName, "-template")
  repo_create(ghorg, repoName, gitignore_template = "")
  
  # If we have big data files, can't use template repos.
  if (use_template_repo)
    repo_set_template(paste0(ghorg, "/", repoName))
  
  message("Pushing local repo")
  system2("./make-repo.sh", c(repoName, assignmentTemplatePath))
  message(glue("Pushed repo to https:://github.com/{ghorg}/{repoName}"))
}
```

```{r, eval=FALSE}
createAssignment("hw/hw04-other-tools")
```

```python
import subprocess, pathlib
file_list = subprocess.run(
  ["git", "ls-files", "-z"], stdout=subprocess.PIPE, cwd=str(pathlib.Path("hw/hw03-agg"))
).stdout.split(b'\0')[:-1]
for f in file_list:
  f = f.decode('utf-8')
  if "solution" in f:
    continue
  print(f)
  
```


# Push repo to students


```{r solo-assignment, eval=FALSE}
pushSoloAssignment <- function(assignmentName, includeStaff = TRUE, usernames = NULL) {
  if (is.null(usernames)) {
    usernames <- roster$github
    if (includeStaff)
      usernames <- c(usernames, "kcarnold", "kimyejae")
  }
  org_create_assignment(
    org = ghorg,
    user = usernames,
    repo = paste0(assignmentName, '-', usernames),
    source_repo = paste0(ghorg, '/', assignmentName, '-template')
  )
}
pushSoloAssignment('hw04')
```


```{r team-assignment eval=FALSE}
lab04_repodata <- roster_withteams %>%
  select(github, team) %>% 
  bind_rows(tibble(github = c("kcarnold", "kimyejae"), team = "lab04-staff")) %>% 
  mutate(
    repo_base = team,
    repo_full = paste0(ghorg, "/", repo_base)
  )

lab04_repodata %>% 
  {
    org_create_assignment(
      org = ghorg,
      user = .$github,
      repo = .$repo_base,
      team = .$team,
      source_repo = paste0(ghorg, "/", "lab04-template")
    )
  }
# org_create_assignment(
#   org = ghorg,
#   user = users,
#   repo = paste0("hw03-", roster$github_username),
#   source_repo = paste0(ghorg, "/", "hw03-template")
# )
```


# Harvest Student Work

```{r}
all_repos <- org_repos(ghorg)
```

Maintain local clones of all student repos:

```{r}
assignment_to_update = "lab04"
local_repo_status <- tibble(repo_full = all_repos) %>% 
  separate(repo_full, c(NA, "repo"), "/", remove = FALSE) %>% 
  mutate(
    repo_url = glue("https://github.com/{repo_full}"),
    local_path = fs::path_expand(file.path("../fa20/all_repos", repo)) %>% fs::path_abs() %>% as.character(),
    exists = file.exists(local_path)
  )
local_repo_status %>%
  filter(!exists) %>% 
  select(repo_url, local_path) %>% 
  pwalk(function(repo_url, local_path) 
    purrr::safely(gert::git_clone)(url = repo_url, path = local_path, branch = "master", verbose = TRUE))
updates <- local_repo_status %>% 
  filter(exists, str_detect(repo, assignment_to_update)) %>%
  pull(local_path) %>% 
  map(local_repo_pull)
```

Get all students who are on all repos

```{r get-collaborators}
repo_to_user <-
  all_repos %>% 
  map_dfr(~ repo_collaborators(.x, include_admins = FALSE))
```

```{r lookup-names}
local_repo_owners <- full_join(
  local_repo_status,
  repo_to_user,
  by = c("repo_full" = "repo")
) %>% left_join(
  roster,
  by = c("username" = "github")
) %>%
  select(repo, local_path, name = preferred_full_name)
```

Make links to each student's work.

```{r}
local_repo_owners %>%
  drop_na() %>% 
  pwalk(function(repo, local_path, name) {
    user_path <- fs::path_wd(paste0("../fa20/", name))
    fs::dir_create(user_path)
    fs::link_create(
      path = local_path,
      new_path = file.path(user_path, repo)
    )
  })
```

Make a table for everyone's work

```{r results='asis'}
assignmentName <- "hw02"

students %>% 
  mutate(
    repo_name = glue("{ghorg}/{assignmentName}-{github_username}"),
    repo_url = glue("https://github.com/{repo_name}"),
    md = glue("* [{name}]({repo_url})")
  ) %>% pull(md) %>% cat(sep = "\n")
```

