---
title: "Repo Management"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# remotes::install_github("rundel/ghclass")
library(ghclass)
library(tidyverse)
library(glue)
#library(R.cache)
```

# Setup

Create a GitHub org, and a GitHub Classroom within that org.

The first exercise I created as a GitHub classroom assignment, so students could accept it using the 
GitHub Classroom machinery. After that, we'll create a Portfolio repo and just manage it ourselves.

```{r}
ghorg <- "Calvin-DS202-21FA"
org_sitrep(ghorg)
```

```{r}
clones_path <- here::here("student_data/repos/")
portfolio_clone_path <- fs::path_join(c(clones_path, "portfolio"))
```

Students who join a GitHub Classroom by accepting an assignment are added as "outside collaborators".
I think this works ok.

I set the TA/grader as an Owner on the GitHub org.

```{r}
org_outside_collaborators <- function(org) {
  ghclass:::ghclass_api_v3_req(endpoint = "GET /orgs/:org/outside_collaborators", org = org) %>%
    purrr::map_chr("login")
}
members <- org_members(ghorg)
students <- members %>% setdiff(org_admins(ghorg))
students <- union(students, org_outside_collaborators(ghorg))
```

The first quiz includes a question for GitHub username. This establishes a clear link between GitHub and Moodle.

```{r}
roster <-
  read_csv(
    "../student_data/21FA DATA-202-A-Quiz 1-responses.csv",
    col_types = cols(.default = col_character())
  ) %>%
  janitor::clean_names() %>%
  select(-(state:grade_100_00)) %>% # remove Moodle stuff
  rename(
    preferred_name = response_1,
    github = response_2,
    major = response_3,
    covid = response_4,
    secret_word = response_5,
    bool_expr = response_6,
    courses = response_7,
    ideas = response_8,
    other = response_9
  )
nrow(roster)

# All org members should be on the roster.
students %>% setdiff(roster$github)

# All roster members should be in the org.
setdiff(roster$github, students)
```

```{r}
roster %>% count(covid, sort = TRUE)# %>% clipr::write_clip()
```

```{r}
roster %>% mutate(
  num_courses = str_count(courses, "[0-9][0-9][0-9]")) %>% 
  ggplot(aes(x = num_courses)) +
    geom_bar() +
    scale_x_continuous(breaks = 1:10) +
    labs(x = "Number of previous courses")
```


# Create portfolio template

Create the repo manually for now. Clone:

```{r}
if (!dir.exists(paste0(clones_path, "/portfolio-reference")))
  local_repo_clone(glue("{ghorg}/portfolio-reference"), clones_path)
```

STOP. Put some files in there by hand. Commit and push by hand.


```{r compute-repo-names}
student_repos <- tibble(
  username = students,
  portfolio_repo = paste0("portfolio-", students),
  portfolio_repo_full = glue("{ghorg}/{portfolio_repo}"),
)
```

IGNORE THE BELOW. Use a GitHub Classroom assignment for this repo also. Because
otherwise students need to accept an emailed invite, and they can lose the email.

Now let's create portfolio repos for everyone.

```{r eval=FALSE}
repo_set_template(paste0(ghorg, "/", "portfolio-reference"))
org_create_assignment(
  org = ghorg,
  user = student_repos$username,
  repo = student_repos$portfolio_repo,
  source_repo = paste0(ghorg, "/", "portfolio-reference")
)
```

# Deploy assignments

Once an assignment is deployed, we should treat that as final. Updating templates in-place can cause merge conflicts for students.

```{r}
# This works because this Rmd file is in the `src` directory, and the directory structure here matches the portfolio repo.
deploy_assignment <- function(path, deploy_solutions = FALSE) {
  # Need to pull out the names so we don't get gobbledygook control characters in the commit messages.
  files <- names(fs::dir_ls(path, recurse = TRUE, type = "file"))
  
  is_solution <- str_detect(files, "solution")
  if (deploy_solutions) {
    files <- files[is_solution]
  } else {
    files <- files[!is_solution]
  }
  
  # Add to the reference
  repo_add_file(glue("{ghorg}/portfolio-reference"), files, preserve_path = TRUE, overwrite = FALSE)

  # Add to the student repos.
  repo_add_file(student_repos$portfolio_repo_full, files, preserve_path = TRUE, overwrite = FALSE)

}
#deploy_assignment("ex/ex02-plotting")
```

