---
title: "Repo Management"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# remotes::install_github("rundel/ghclass")
library(ghclass)
library(tidyverse)
library(glue)
#library(R.cache)
```

# Setup

Create a GitHub org, and a GitHub Classroom within that org.

The first exercise I created as a GitHub classroom assignment, so students could accept it using the 
GitHub Classroom machinery. After that, we'll create a Portfolio repo and just manage it ourselves.

```{r}
ghorg <- "Calvin-DS202-21FA"
org_sitrep(ghorg)
```

```{r}
clones_path <- here::here("student_data/repos/")
portfolio_clone_path <- fs::path_join(c(clones_path, "portfolio"))
```

Students who join a GitHub Classroom by accepting an assignment are added as "outside collaborators".
I think this works ok.

I set the TA/grader as an Owner on the GitHub org.

```{r}
org_outside_collaborators <- function(org) {
  ghclass:::ghclass_api_v3_req(endpoint = "GET /orgs/:org/outside_collaborators", org = org) %>%
    purrr::map_chr("login")
}
members <- org_members(ghorg)
students <- members %>% setdiff(org_admins(ghorg))
students <- union(students, org_outside_collaborators(ghorg))
```

The first quiz includes a question for GitHub username. This establishes a clear link between GitHub and Moodle.

```{r}
roster <- read_csv("../student_data/21FA DATA-202-A-Quiz 1-responses.csv", col_types = cols(.default = col_character())) %>%  
  janitor::clean_names() %>% 
  select(-(state:grade_100_00)) %>% # remove Moodle stuff
  rename(preferred_name = response_1, github = response_2, major = response_3, covid = response_4, secret_word = response_5,
         bool_expr = response_6, courses=response_7, ideas=response_8, other=response_9)
nrow(roster)

# All org members should be on the roster.
students %>% setdiff(roster$github)

# All roster members should be in the org.
setdiff(roster$github, students)
```

```{r}
roster %>% count(covid, sort = TRUE)
```


# Create portfolio template

Create the repo manually for now. Clone:

```{r}
local_repo_clone(glue("{ghorg}/portfolio-reference"), clones_path)
```

STOP. Put some files in there by hand. Commit and push by hand.

Now let's create portfolio repos for everyone.

```{r eval=FALSE}
repo_set_template(paste0(ghorg, "/", "portfolio-reference"))
org_create_assignment(
  org = ghorg,
  user = students,
  repo = paste0("portfolio-", students),
  source_repo = paste0(ghorg, "/", "portfolio-reference")
)
```

