---
title: "Wrangling practice: one table"
author: "DATA 202 21FA, based on [datasciencebox.org](https://datasciencebox.org/)"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "4:3"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "../slide-setup.Rmd"}
```

```{r packages, include=FALSE}
library(dplyr)
library(readr)
library(ggplot2)
```


## Q&A

> Will we be working with different data?

We'll use a mix of old and new examples each week.

> When should we spend effort making graphs look good?

Homework, exams: always. Otherwise not necessary but always good practice.

> How do we remember which types of plots to make and how?

Class slides, RStudio cheatsheets, textbook, ...

---

class: center, middle

## Today: Practice thinking about grammar of transformation

Friday: in-lab practice

---

## Data: Hotel bookings

- Data from two hotels: one resort and one city hotel
- Observations: Each row represents a hotel booking
- Goal for original data collection: Development of prediction models to classify a hotel booking's likelihood to be cancelled ([Antonia et al., 2019](https://www.sciencedirect.com/science/article/pii/S2352340918315191#bib5))

```{r message=FALSE}
hotels <- read_csv("data/hotels.csv")
head(hotels)
#head(select(hotels, hotel, is_canceled, lead_time), 5)
```

.floating-source[
Source: [TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md)
]


---
layout: true

## ascending / descending order

---

.pull-left[
```{r}
hotels %>%
  select(adults, children, babies) %>%
  arrange(babies) #<<
```
]

--

.pull-right[
```{r}
hotels %>%
  select(adults, children, babies) %>%
  arrange(desc(babies)) #<<
```
]

---

.pull-left[
```{r}
hotels %>%
  select(adults, children, babies) %>%
  arrange(babies) #<<
```
]
.pull-right[
```{r}
hotels %>%
  select(adults, children, babies) %>%
  arrange(-babies) #<<
```
]

---
layout: true

## `filter`

How could we remove the resort hotel? (`hotel` being `"Resort Hotel"`)

---

```{r eval=FALSE}
hotels %>%
  filter(___)
```

---

```{r}
hotels %>% 
  filter(hotel != "Resort Hotel")
```

---
layout: false

## `filter` keeps rows matching conditions

---

## Multiple conditions

Any bookings for only kids (no `adults`, but `children` or `babies`)?

--

```{r}
hotels %>%
  filter( 
    adults == 0,     
    children >= 1 | babies >= 1     # | means or  #<<
    ) %>%
  select(adults, babies, children)
```

---

## Logical operators in R

<br>

operator    | definition                   || operator     | definition
------------|------------------------------||--------------|----------------
`<`         | less than                    ||`x`&nbsp;&#124;&nbsp;`y`     | `x` OR `y` 
`<=`        |	less than or equal to        ||`is.na(x)`    | test if `x` is `NA`
`>`         | greater than                 ||`!is.na(x)`   | test if `x` is not `NA`
`>=`        |	greater than or equal to     ||`x %in% y`    | test if `x` is in `y`
`==`        |	exactly equal to             ||`!(x %in% y)` | test if `x` is not in `y`
`!=`        |	not equal to                 ||`!x`          | not `x`
`x & y`     | `x` AND `y`                  ||              |

---
layout: true

## Which market segments book most in advance?

.pull-left-narrow.smaller-table[
```{r echo=FALSE}
hotels %>% 
  group_by(market_segment) %>% 
  summarize(lead_time = mean(lead_time)) %>% 
  arrange(desc(lead_time)) %>% 
  knitr::kable()
```
]

---

.pull-right-wide[
How would we compute this?
]

---

.pull-right-wide[
```{r}
hotels %>% 
  group_by(market_segment) %>% 
  summarize(
    lead_time = mean(lead_time))
```
]

---

.pull-right-wide[
```{r}
hotels %>% 
  group_by(market_segment) %>% 
  summarize(
    lead_time = mean(lead_time)) %>% 
  arrange(desc(lead_time))
```
]

---
layout: true

## How many total nights stayed?

---

```{r}
hotels %>% 
  select(hotel, stays_in_week_nights, stays_in_weekend_nights)
```

```{r eval=FALSE}
hotels %>% 
  mutate( #<<
    num_nights = ____)
```

---

```{r}
hotels %>%
  mutate( #<<
    num_nights = stays_in_week_nights + stays_in_weekend_nights) %>% 
  select(hotel, num_nights)
```

---
layout: false

## How long did each market segment stay?

--

```{r}
hotels %>%
  mutate(
    num_nights = stays_in_week_nights + stays_in_weekend_nights) %>% 
  group_by(market_segment) %>% 
  summarize(num_nights = mean(num_nights)) %>% 
  arrange(desc(num_nights))
```

